---
title: "Hindcast ROMS variables"
author: "Rebecca Howard"
date: "9/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(marmap)
library(maps)
library(mapdata)
library(fields)
library(here)
library(mgcv)
library(tidyr)
library(dplyr)
library(ggplot2)
library(raster)
library(rgdal)
library(viridis)
library(RColorBrewer)
source(here('code/functions', 'vis_gam_COLORS.R'))
source(here('code/functions', 'distance_function.R'))
source(here('code/functions', 'gam_selection.R'))
source(here('code/functions', 'map_vc_presentation.R'))
source(here('code/functions', 'map_vc.R'))
source(here('code/functions', 'map_phenology.R'))
source(here('code/functions', 'threepanel_base_mse.R'))
source(here('code/functions', 'vc_grids_presentation.R'))
source(here('code/functions', 'vc_grids.R'))
source(here('code/functions', 'plot_variable.R'))
source(here('code/functions', 'location_plot.R'))
```

```{r}
# Load ROMS temperature means
roms_temps <- readRDS(here('data', 'roms_temps.rds'))

# Load fish data
clean_data <- function(file, temperature){
df <- as.data.frame(filter((readRDS(here('data', file))),
                 lat >= 52 & lat <= 62,
                  lon >= -176.5 & lon <= -156.5))
df$mean_temp <- temperature$mean[match(df$year, temperature$year)]
df$catch <- df$larvalcatchper10m2 + 1
df <- na.omit(df)
return(df)
}

pk_egg <- clean_data('pk_egg.rds', roms_temps)
pk_larvae <- clean_data('pk_larvae.rds', roms_temps)
fhs_egg <- clean_data('fhs_egg.rds', roms_temps)
fhs_larvae <- clean_data('fhs_larvae.rds', roms_temps)
akp_egg <- clean_data('akp_egg.rds', roms_temps)
akp_larvae <- clean_data('akp_larvae.rds', roms_temps)
yfs_egg <- clean_data('yfs_egg.rds', roms_temps)
yfs_larvae <- clean_data('yfs_larvae.rds', roms_temps)
pcod_larvae <- clean_data('pcod_larvae.rds', roms_temps)
nrs_larvae <- clean_data('nrs_larvae.rds', roms_temps)

# Colors
contour_col <- rgb(0, 0, 255, max = 255, alpha = 0, names = "white")

jet.colors = colorRampPalette(rev(c("#FFFFCC", "#FBF2A8", "#F9E585",
                                    "#F5D363", "#EFBA55", "#EAA352",
                                    "#E68C51", "#E0754F", "#D75C4D",
                                    "#BB4A48", "#994240", "#763931", 
                                    "#542D20", "#352311", "#191900")))
```

## Pollock
### Tweedie
Base models
```{r}
# Tweedie
pk_egg_base_t <- gam(catch ~ factor(year) +
                       s(lon, lat) + 
                       s(doy),
                    data = pk_egg,
                    family = tw(link = 'log'),
                    method = 'REML')
summary(pk_egg_base_t)

pk_larvae_base_t <- gam(catch ~ factor(year) +
                       s(lon, lat) + 
                       s(doy),
                    data = pk_larvae,
                    family = tw(link = 'log'),
                    method = 'REML')
summary(pk_larvae_base_t)
```
#### Eggs

Models with flexible phenology or geography 
Note variable coefficient by year for each
This takes forever to run, don't do unless necessary to recreate figure
```{r}
pk_egg_phenology_t <- gam(catch ~ factor(year) +
                            s(doy) +
                            s(lon, lat) +
                            s(doy, by = factor(year)),
                          data = pk_egg,
                          family = tw(link = 'log'),
                          method =  'REML')
summary(pk_egg_phenology_t)
AIC(pk_egg_phenology_t)
# AIC: 33996

pk_egg_geography_t <- gam(catch ~ factor(year) +
                            s(lon, lat) +
                            s(lon, lat, by = factor(year)) +
                            s(doy),
                          data = pk_egg,
                          family = tw(link = 'log'),
                          method = 'REML')
summary(pk_egg_geography_t)
AIC(pk_egg_geography_t)
# AIC: 33177
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_pkegg_phenology_t <- (summary(pk_egg_base_t)$scale - 
                      summary(pk_egg_phenology_t)$scale) / summary(pk_egg_base_t)$scale
ratio_pkegg_phenology_t #0.111

ratio_pkegg_geography_t <- (summary(pk_egg_base_t)$scale - 
                      summary(pk_egg_geography_t)$scale) / summary(pk_egg_base_t)$scale
ratio_pkegg_geography_t #0.310
```

Add ROMS mean temperatures Use a representative area with averaged values from Feb - April for each year This can then be matched to the egg/larvae data for use in variable coefficient by SST instead of year as above

Determine best GAM from 7 formulations (see function)
```{r}
pkegg_gams_t <- gam_selection(pk_egg)
AIC(pkegg_gams_t[[1]][[1]]) 
AIC(pkegg_gams_t[[1]][[2]]) 
AIC(pkegg_gams_t[[1]][[3]]) 
summary(pkegg_gams_t[[2]]) # best model is temporal VC, retain all smooth terms
AIC(pkegg_gams_t[[1]][[1]]) - AIC(pkegg_gams_t[[1]][[3]]) # 63.6
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(pkegg_gams_t[[1]][[2]], select = 1, main = "DOY")
plot(pkegg_gams_t[[1]][[2]], select = 3, main = "Temperature")
plot(pkegg_gams_t[[1]][[2]], select = 4, main = "Salinity")
plot(pkegg_gams_t[[1]][[2]], select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_egg_roms_monthvc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(pkegg_gams_t[[1]][[2]])
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_egg_roms_monthvc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(pkegg_gams_t[[1]][[3]], select = 1, main = "DOY")
plot(pkegg_gams_t[[1]][[3]], select = 3, main = "Temperature")
plot(pkegg_gams_t[[1]][[3]], select = 4, main = "Salinity")
plot(pkegg_gams_t[[1]][[3]], select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_egg_roms_spacevc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(pkegg_gams_t[[1]][[3]])
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_egg_roms_spacevc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_pkegg_month_t <- (summary(pk_egg_base_t)$scale - 
                         summary(pkegg_gams_t[[1]][[2]])$scale) / summary(pk_egg_base_t)$scale
var_ratio_pkegg_month_t #0.0418

var_ratio_pkegg_space_t <- (summary(pk_egg_base_t)$scale - 
                         summary(pkegg_gams_t[[1]][[3]])$scale) / summary(pk_egg_base_t)$scale
var_ratio_pkegg_space_t #0.0562
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
pkegg_grids_t <- threepanel_base_mse(pk_egg, pk_egg_base_t)
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(pk_egg, pkegg_grids_t, ratio_pkegg_geography_t,
                 var_ratio_pkegg_space_t,
                 ratio_pkegg_phenology_t,
                 var_ratio_pkegg_month_t)
dev.copy(jpeg,
         here('results/pollock_hindcast', 
              'pollock_egg_MSE_t.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
```{r}
pkegg_vcgrids_t <- vc_grids_presentation(pk_egg, pkegg_gams_t)
```

Create two panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc_presentation(pk_egg, pkegg_gams_t[[2]], pkegg_vcgrids_t, "Hindcast")
legend(50,
       6,
       legend = c('Mean SST', '+ 1C'),
       col = c('#000000', '#E69F00'),
       lty = 1,
       bty = 'n',
       lwd = 2,
       cex = 1.2)
dev.copy(jpeg,
         here('results/pollock_hindcast',
         'pollock_egg_vc_t.jpg'),
         height = 3.5,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

```{r}
tiff(here('results/pollock_hindcast/outputs',
         'pollock_egg_output.jpg'),
     units = "in",
     width = 40,
     height = 12,
     res = 200)
par(mfrow = c(1, 3),
    mar = c(11, 15, .5, 0.6) + 0.1,
    oma = c(3, 1, 1, 1),
    mgp = c(9, 4, 0))
plot_variable(pkegg_gams_t,
              covariate = 1,
              bounds = c(-5.5, 7),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(pkegg_gams_t,
              covariate = 3,
              bounds = c(-5.5, 7),
              "Temperature",
              " ",
              "n")
plot_variable(pkegg_gams_t,
              covariate = 4,
              bounds = c(-5.5, 7),
              "Salinity",
              " ",
              "n")

dev.off()
```

Plot the variable coefficient term
```{r}
pk_egg <- variable_coefficient(pkegg_gams_t, pk_egg)

vis.gam(pkegg_gams_t[[2]],
        view = c('lon', 'lat'),
        plot.type = "contour",
        color = "gray",
        too.far = 0.07,
        type = 'response',
        main = 'All Years - All temps',
        ylab = 'Latitude',
        xlab = 'Longitude')
# symbols(pk_egg$lon[pk_egg$sign_slope_pos],
#         pk_egg$lat[pk_egg$sign_slope_pos],
#         circle = pk_egg$pred_slope[pk_egg$sign_slope_pos],
#         inches = 0.08,
#         add = T,
#         bg = alpha('darkred', 0.4),
#         fg = alpha('black', 0.08))
symbols(pk_egg$lon[pk_egg$sign_slope_neg],
        pk_egg$lat[pk_egg$sign_slope_neg],
        circle = (-1) * pk_egg$pred_slope[pk_egg$sign_slope_neg],
        inches = 0.08,
        add = T,
        bg = alpha('darkslateblue', 0.4),
        fg = alpha('black', 0.08))
map("worldHires",
    fill = T,
    col = "wheat4",
    add = T)
dev.copy(jpeg,
         here('results/pollock_hindcast',
         'pollock_egg_vcterm.jpg'),
         height = 10,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

#### Larvae

Models with flexible phenology or geography 
Note variable coefficient by year for each
```{r}
pk_larvae_phenology_t <- gam(catch ~ factor(year) +
                               s(doy) +
                               s(lon, lat) +
                               s(doy, by = factor(year)),
                             data = pk_larvae,
                             family = tw(link = 'log'),
                             method = 'REML')
summary(pk_larvae_phenology_t)
AIC(pk_larvae_phenology_t)
# AIC: 32891

pk_larvae_geography_t <- gam(catch ~ factor(year) +
                               s(lon, lat) +
                               s(lon, lat, by = factor(year)) +
                               s(doy),
                             data = pk_larvae,
                             family = tw(link = 'log'),
                             method = 'REML')
summary(pk_larvae_geography_t)
AIC(pk_larvae_geography_t)
# AIC: 31370
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_pklarvae_phenology_t <- (summary(pk_larvae_base_t)$scale - 
                      summary(pk_larvae_phenology_t)$scale) / summary(pk_larvae_base_t)$scale
ratio_pklarvae_phenology_t # 0.119

ratio_pklarvae_geography_t <- (summary(pk_larvae_base_t)$scale - 
                      summary(pk_larvae_geography_t)$scale) / summary(pk_larvae_base_t)$scale
ratio_pklarvae_geography_t # 0.426
```

Determine best GAM from 7 formulations (see function)
```{r}
pklarvae_gams_t <- gam_selection(pk_larvae)
AIC(pklarvae_gams_t[[1]][[1]]) #33345
AIC(pklarvae_gams_t[[1]][[2]]) #33102
AIC(pklarvae_gams_t[[1]][[3]]) #33238
summary(pklarvae_gams_t[[2]]) # best model is temporal, retain all smooth terms
AIC(pklarvae_gams_t[[1]][[1]]) - AIC(pklarvae_gams_t[[2]]) #242.8
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(pklarvae_gams_t[[1]][[2]], select = 1, main = "DOY")
plot(pklarvae_gams_t[[1]][[2]], select = 3, main = "Temperature")
plot(pklarvae_gams_t[[1]][[2]], select = 4, main = "Salinity")
plot(pklarvae_gams_t[[1]][[2]], select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_larvae_roms_monthvc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(pklarvae_gams_t[[1]][[2]])
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_larvae_roms_monthvc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(pklarvae_gams_t[[1]][[3]], select = 1, main = "DOY")
plot(pklarvae_gams_t[[1]][[3]], select = 3, main = "Temperature")
plot(pklarvae_gams_t[[1]][[3]], select = 4, main = "Salinity")
plot(pklarvae_gams_t[[1]][[3]], select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_larvae_roms_spacevc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(pklarvae_gams_t[[1]][[3]])
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_larvae_roms_spacevc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_pklarvae_month_t <- (summary(pk_larvae_base_t)$scale - 
                         summary(pklarvae_gams_t[[1]][[2]])$scale) / summary(pk_larvae_base_t)$scale
var_ratio_pklarvae_month_t #0.0737

var_ratio_pklarvae_space_t <- (summary(pk_larvae_base_t)$scale - 
                         summary(pklarvae_gams_t[[1]][[3]])$scale) / summary(pk_larvae_base_t)$scale
var_ratio_pklarvae_space_t #0.0900
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
pklarvae_grids_t <- threepanel_base_mse(pk_larvae, pk_larvae_base_t)
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(pk_larvae, pklarvae_grids_t, ratio_pklarvae_geography_t,
                 var_ratio_pklarvae_space_t,
                 ratio_pklarvae_phenology_t,
                 var_ratio_pklarvae_month_t)
dev.copy(jpeg,
         here('results/pollock_hindcast', 
              'pollock_larvae_MSE_t.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
```{r}
pklarvae_vcgrids_t <- vc_grids_presentation(pk_larvae, pklarvae_gams_t)
```

Create two panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc_presentation(pk_larvae, pklarvae_gams_t[[2]], pklarvae_vcgrids_t, "Spatial GAM hindcast")
legend(50,
       2,
       legend = c('Mean SST', '+ 1C'),
       col = c('#000000', '#E69F00'),
       lty = 1,
       bty = 'n',
       lwd = 2,
       cex = 1.2)
dev.copy(jpeg,
         here('results/pollock_hindcast',
         'pollock_larvae_vc_t.jpg'),
         height = 3.5,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

```{r}
tiff(here('results/pollock_hindcast/outputs',
         'pollock_larvae_output.jpg'),
     units = "in",
     width = 40,
     height = 12,
     res = 200)

par(mfrow = c(1, 3),
    mar = c(11, 15, .5, 0.6) + 0.1,
    oma = c(3, 1, 1, 1),
    mgp = c(9, 4, 0))
plot_variable(pklarvae_gams_t,
              covariate = 1,
              bounds = c(-5.5, 7),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(pklarvae_gams_t,
              covariate = 3,
              bounds = c(-5.5, 7),
              "Temperature",
              " ",
              "n")
plot_variable(pklarvae_gams_t,
              covariate = 4,
              bounds = c(-5.5, 7),
              "Salinity",
              " ",
              "n")

dev.off()
```

Plot the variable coefficient term
```{r}
pk_larvae <- variable_coefficient(pklarvae_gams_t, pk_larvae)

vis.gam(pklarvae_gams_t[[2]],
        view = c('lon', 'lat'),
        plot.type = "contour",
        color = "gray",
        too.far = 0.07,
        type = 'response',
        main = 'All Years - All temps',
        ylab = 'Latitude',
        xlab = 'Longitude')
# symbols(pk_larvae$lon[pk_larvae$sign_slope_pos],
#         pk_larvae$lat[pk_larvae$sign_slope_pos],
#         circle = pk_larvae$pred_slope[pk_larvae$sign_slope_pos],
#         inches = 0.08,
#         add = T,
#         bg = alpha('darkred', 0.4),
#         fg = alpha('black', 0.08))
symbols(pk_larvae$lon[pk_larvae$sign_slope_neg],
        pk_larvae$lat[pk_larvae$sign_slope_neg],
        circle = (-1) * pk_larvae$pred_slope[pk_larvae$sign_slope_neg],
        inches = 0.08,
        add = T,
        bg = alpha('darkslateblue', 0.4),
        fg = alpha('black', 0.08))
map("worldHires",
    fill = T,
    col = "wheat4",
    add = T)
dev.copy(jpeg,
         here('results/pollock_hindcast',
         'pollock_larvae_vcterm.jpg'),
         height = 10,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

## Flathead Sole
### Tweedie
```{r}
fhs_egg_base_t <- gam(catch ~ factor(year) +
                        s(lon, lat) +
                        s(doy),
                      data = fhs_egg,
                      family = tw(link = 'log'),
                      method = 'REML')
summary(fhs_egg_base_t)

fhs_larvae_base_t <- gam(catch ~ factor(year) +
                           s(lon, lat) +
                           s(doy),
                         data = fhs_larvae,
                         family = tw(link = 'log'),
                         method = 'REML')
summary(fhs_larvae_base_t)
```

#### Eggs

Models with flexible phenology or geography 
Note variable coefficient by year for each
This takes forever to run, don't do unless necessary to recreate figure
```{r}
fhs_egg_phenology_t <- gam(catch ~ factor(year) +
                             s(doy) +
                             s(lon, lat) +
                             s(doy, by = factor(year)),
                           data = fhs_egg,
                           family = tw(link = 'log'),
                           method = 'REML')
summary(fhs_egg_phenology_t)
AIC(fhs_egg_phenology_t)
# AIC: 9907

fhs_egg_geography_t <- gam(catch ~ factor(year) +
                             s(lon, lat) +
                             s(lon, lat, by = factor(year)) +
                             s(doy),
                           data = fhs_egg,
                           family = tw(link = 'log'),
                           method =  'REML')
summary(fhs_egg_geography_t)
AIC(fhs_egg_geography_t)
# AIC: 8939
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_fhsegg_phenology_t <- (summary(fhs_egg_base_t)$scale - 
                      summary(fhs_egg_phenology_t)$scale) / summary(fhs_egg_base_t)$scale
ratio_fhsegg_phenology_t #0.171

ratio_fhsegg_geography_t <- (summary(fhs_egg_base_t)$scale - 
                      summary(fhs_egg_geography_t)$scale) / summary(fhs_egg_base_t)$scale
ratio_fhsegg_geography_t #0.496
```

Add ROMS mean temperatures Use a representative area with averaged values from Feb - April for each year This can then be matched to the egg/larvae data for use in variable coefficient by SST instead of year as above


Determine best GAM from 7 formulations (see function)
```{r}
fhsegg_gams_t <- gam_selection(fhs_egg)
AIC(fhsegg_gams_t[[1]][[1]]) # 13559
AIC(fhsegg_gams_t[[1]][[2]]) # 13343
AIC(fhsegg_gams_t[[1]][[3]]) # 13536
summary(fhsegg_gams_t[[2]]) # best model is doy VC, remove salinity
AIC(fhsegg_gams_t[[1]][[1]]) - AIC(fhsegg_gams_t[[2]]) #216.7
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(fhsegg_gams_t[[1]][[2]], select = 1, main = "DOY")
plot(fhsegg_gams_t[[1]][[2]], select = 3, main = "Temperature")
plot(fhsegg_gams_t[[1]][[2]], select = 4, main = "Salinity")
plot(fhsegg_gams_t[[1]][[2]], select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/flathead_hindcast', 'flathead_egg_roms_monthvc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(fhsegg_gams_t[[1]][[2]])
dev.copy(jpeg, here('results/flathead_hindcast', 'flathead_egg_roms_monthvc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(fhsegg_gams_t[[1]][[3]], select = 1, main = "DOY")
plot(fhsegg_gams_t[[1]][[3]], select = 3, main = "Temperature")
plot(fhsegg_gams_t[[1]][[3]], select = 4, main = "Salinity")
plot(fhsegg_gams_t[[1]][[3]], select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/flathead_hindcast', 'flathead_egg_roms_spacevc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(fhsegg_gams_t[[1]][[3]])
dev.copy(jpeg, here('results/flathead_hindcast', 'flathead_egg_roms_spacevc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_fhsegg_month_t <- (summary(fhs_egg_base_t)$scale - 
                         summary(fhsegg_gams_t[[1]][[2]])$scale) / summary(fhs_egg_base_t)$scale
var_ratio_fhsegg_month_t #0.230

var_ratio_fhsegg_space_t <- (summary(fhs_egg_base_t)$scale - 
                         summary(fhsegg_gams_t[[1]][[3]])$scale) / summary(fhs_egg_base_t)$scale
var_ratio_fhsegg_space_t #0.189
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
fhsegg_grids_t <- threepanel_base_mse(fhs_egg, fhs_egg_base_t)
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(fhs_egg, fhsegg_grids_t, ratio_fhsegg_geography_t,
                 var_ratio_fhsegg_space_t,
                 ratio_fhsegg_phenology_t,
                 var_ratio_fhsegg_month_t)
dev.copy(jpeg,
         here('results/flathead_hindcast', 
              'flathead_egg_MSE_t.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
```{r}
fhsegg_vcgrids_t <- vc_grids_presentation(fhs_egg, fhsegg_gams_t)
```

Create three panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc_presentation(fhs_egg, fhsegg_gams_t[[2]], fhsegg_vcgrids_t, "Spatial GAM hindcast")
legend(130,
       1,
       legend = c('Mean SST', '+ 1C'),
       col = c('#000000', '#E69F00'),
       lty = 1,
       bty = 'n',
       lwd = 2,
       cex = 1.2)
dev.copy(jpeg,
         here('results/flathead_hindcast',
         'flathead_egg_vc_t.jpg'),
         height = 3.5,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

```{r}
tiff(here('results/flathead_hindcast/outputs',
         'flathead_egg_output.jpg'),
     units = "in",
     width = 40,
     height = 12,
     res = 200)

par(mfrow = c(1, 3),
    mar = c(11, 15, .5, 0.6) + 0.1,
    oma = c(3, 1, 1, 1),
    mgp = c(9, 4, 0))
plot_variable(fhsegg_gams_t,
              covariate = 1,
              bounds = c(-5, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(fhsegg_gams_t,
              covariate = 3,
              bounds = c(-5, 3),
              "Temperature",
              " ",
              "n")
plot_variable(fhsegg_gams_t,
              covariate = 4,
              bounds = c(-5, 3),
              "Salinity",
              " ",
              "n")

dev.off()
```


#### Larvae

Models with flexible phenology or geography 
Note variable coefficient by year for each
```{r}
fhs_larvae_phenology_t <- gam(catch ~ factor(year) +
                                s(doy) +
                                s(lon, lat) +
                                s(doy, by = factor(year)),
                              data = fhs_larvae,
                              family = tw(link = 'log'),
                              method = 'REML')
summary(fhs_larvae_phenology_t)
AIC(fhs_larvae_phenology_t)
# AIC: 22747

fhs_larvae_geography_t <- gam(catch ~ factor(year) +
                                s(lon, lat) +
                                s(lon, lat, by = factor(year)) +
                                s(doy),
                              data = fhs_larvae,
                              family = tw(link = 'log'),
                              method = 'REML')
summary(fhs_larvae_geography_t)
AIC(fhs_larvae_geography_t)
# AIC: 21956
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_fhslarvae_phenology_t <- (summary(fhs_larvae_base_t)$scale - 
                      summary(fhs_larvae_phenology_t)$scale) / summary(fhs_larvae_base_t)$scale
ratio_fhslarvae_phenology_t # 0.121

ratio_fhslarvae_geography_t <- (summary(fhs_larvae_base_t)$scale - 
                      summary(fhs_larvae_geography_t)$scale) / summary(fhs_larvae_base_t)$scale
ratio_fhslarvae_geography_t # 0.271
```

Add ROMS mean temperatures Use a representative area with averaged values from Feb - April for each year This can then be matched to the larvae/larvae data for use in variable coefficient by SST instead of year as above

Determine best GAM from 7 formulations (see function)
```{r}
fhslarvae_gams_t <- gam_selection(fhs_larvae)
AIC(fhslarvae_gams_t[[1]][[1]]) #14864
AIC(fhslarvae_gams_t[[1]][[2]]) #14847
AIC(fhslarvae_gams_t[[1]][[3]]) #14843
summary(fhslarvae_gams_t[[2]]) # best model is spatial VC, keep all smooth terms
AIC(fhslarvae_gams_t[[1]][[1]]) - AIC(fhslarvae_gams_t[[2]]) #20.2
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(fhslarvae_gams_t[[1]][[2]], select = 1, main = "DOY")
plot(fhslarvae_gams_t[[1]][[2]], select = 3, main = "Temperature")
plot(fhslarvae_gams_t[[1]][[2]], select = 4, main = "Salinity")
plot(fhslarvae_gams_t[[1]][[2]], select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/flathead_hindcast', 'flathead_larvae_roms_monthvc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(fhslarvae_gams_t[[1]][[2]])
dev.copy(jpeg, here('results/flathead_hindcast', 'flathead_larvae_roms_monthvc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(fhslarvae_gams_t[[1]][[3]], select = 1, main = "DOY")
plot(fhslarvae_gams_t[[1]][[3]], select = 3, main = "Temperature")
plot(fhslarvae_gams_t[[1]][[3]], select = 4, main = "Salinity")
plot(fhslarvae_gams_t[[1]][[3]], select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/flathead_hindcast', 'flathead_larvae_roms_spacevc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(fhslarvae_gams_t[[1]][[3]])
dev.copy(jpeg, here('results/flathead_hindcast', 'flathead_larvae_roms_spacevc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_fhslarvae_month_t <- (summary(fhs_larvae_base_t)$scale - 
                         summary(fhslarvae_gams_t[[1]][[2]])$scale) / summary(fhs_larvae_base_t)$scale
var_ratio_fhslarvae_month_t #0.0701

var_ratio_fhslarvae_space_t <- (summary(fhs_larvae_base_t)$scale - 
                         summary(fhslarvae_gams_t[[1]][[3]])$scale) / summary(fhs_larvae_base_t)$scale
var_ratio_fhslarvae_space_t #0.0725
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
fhslarvae_grids_t <- threepanel_base_mse(fhs_larvae, fhs_larvae_base_t)
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(fhs_larvae, fhslarvae_grids_t, ratio_fhslarvae_geography_t,
                 var_ratio_fhslarvae_space_t,
                 ratio_fhslarvae_phenology_t,
                 var_ratio_fhslarvae_month_t)
dev.copy(jpeg,
         here('results/flathead_hindcast', 
              'flathead_larvae_MSE_t.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
```{r}
fhslarvae_vcgrids_t <- vc_grids_presentation(fhs_larvae, fhslarvae_gams_t)
```

Create three panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc_presentation(fhs_larvae, fhslarvae_gams_t[[2]], fhslarvae_vcgrids_t, "Spatial GAM hindcast")
legend(40,
       0.5,
       legend = c('Mean SST', '+ 1C'),
       col = c('#000000', '#E69F00'),
       lty = 1,
       bty = 'n',
       lwd = 2,
       cex = 1.2)
dev.copy(jpeg,
         here('results/flathead_hindcast',
         'flathead_larvae_vc_t.jpg'),
         height = 3.5,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

```{r}
tiff(here('results/flathead_hindcast/outputs',
         'flathead_larvae_output.jpg'),
     units = "in",
     width = 40,
     height = 12,
     res = 200)

par(mfrow = c(1, 3),
    mar = c(11, 15, .5, 0.6) + 0.1,
    oma = c(3, 1, 1, 1),
    mgp = c(9, 4, 0))
plot_variable(fhslarvae_gams_t,
              covariate = 1,
              bounds = c(-5, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(fhslarvae_gams_t,
              covariate = 3,
              bounds = c(-5, 3),
              "Temperature",
              " ",
              "n")
plot_variable(fhslarvae_gams_t,
              covariate = 4,
              bounds = c(-5, 3),
              "Salinity",
              " ",
              "n")
dev.off()
```

## Yellowfin Sole
### Tweedie
```{r}
yfs_egg_base_t <- gam(catch ~ factor(year) +
                        s(lon, lat) +
                        s(doy),
                      data = yfs_egg,
                      family = tw(link = 'log'),
                      method = 'REML')
summary(yfs_egg_base_t)

yfs_larvae_base_t <- gam(catch ~ factor(year) +
                           s(lon, lat) +
                           s(doy),
                         data = yfs_larvae,
                         family = tw(link = 'log'),
                         method = 'REML')
summary(yfs_larvae_base_t)
```

#### Eggs

Models with flexible phenology or geography 
Note variable coefficient by year for each
This takes forever to run, don't do unless necessary to recreate figure
```{r}
yfs_egg_phenology_t <- gam(catch ~ factor(year) +
                             s(doy) +
                             s(lon, lat) +
                             s(doy, by = factor(year)),
                           data = yfs_egg,
                           family = tw(link = 'log'),
                           method = 'REML')
summary(yfs_egg_phenology_t)
AIC(yfs_egg_phenology_t)
# AIC: 1586

yfs_egg_geography_t <- gam(catch ~ factor(year) +
                             s(lon, lat) +
                             s(lon, lat, by = factor(year)) +
                             s(doy),
                           data = yfs_egg,
                           family = tw(link = 'log'),
                           method =  'REML')
summary(yfs_egg_geography_t)
AIC(yfs_egg_geography_t)
# AIC: 997
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_yfsegg_phenology_t <- (summary(yfs_egg_base_t)$scale - 
                      summary(yfs_egg_phenology_t)$scale) / summary(yfs_egg_base_t)$scale
ratio_yfsegg_phenology_t # 0.250

ratio_yfsegg_geography_t <- (summary(yfs_egg_base_t)$scale - 
                      summary(yfs_egg_geography_t)$scale) / summary(yfs_egg_base_t)$scale
ratio_yfsegg_geography_t #0.620
```

Add ROMS mean temperatures Use a representative area with averaged values from Feb - April for each year This can then be matched to the egg/larvae data for use in variable coefficient by SST instead of year as above


Determine best GAM from 7 formulations (see function)
```{r}
yfsegg_gams_t <- gam_selection(yfs_egg)
AIC(yfsegg_gams_t[[1]][[1]]) # 1830
AIC(yfsegg_gams_t[[1]][[2]]) # 1774
AIC(yfsegg_gams_t[[1]][[3]]) # 1774
summary(yfsegg_gams_t[[2]]) # best model is base model, use VC spatial, no environmental smooth terms significant
AIC(yfsegg_gams_t[[1]][[2]]) - AIC(yfsegg_gams_t[[2]]) #0
```

Plot best model for phenology
```{r}
# Best model by AIC is model 1
par(mfrow = c(2, 2))
plot(yfsegg_gams_t[[1]][[2]], select = 1, main = "DOY")
plot(yfsegg_gams_t[[1]][[2]], select = 3, main = "Temperature")
plot(yfsegg_gams_t[[1]][[2]], select = 4, main = "Salinity")
plot(yfsegg_gams_t[[1]][[2]], select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/yellowfin_hindcast', 'yellowfin_egg_roms_monthvc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(yfsegg_gams_t[[1]][[2]])
dev.copy(jpeg, here('results/yellowfin_hindcast', 'yellowfin_egg_roms_monthvc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(yfsegg_gams_t[[1]][[3]], select = 1, main = "DOY")
plot(yfsegg_gams_t[[1]][[3]], select = 3, main = "Temperature")
plot(yfsegg_gams_t[[1]][[3]], select = 4, main = "Salinity")
plot(yfsegg_gams_t[[1]][[3]], select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/yellowfin_hindcast', 'yellowfin_egg_roms_spacevc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(yfsegg_gams_t[[1]][[3]])
dev.copy(jpeg, here('results/yellowfin_hindcast', 'yellowfin_egg_roms_spacevc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_yfsegg_month_t <- (summary(yfs_egg_base_t)$scale - 
                         summary(yfsegg_gams_t[[1]][[2]])$scale) / summary(yfs_egg_base_t)$scale
var_ratio_yfsegg_month_t #-0.0322

var_ratio_yfsegg_space_t <- (summary(yfs_egg_base_t)$scale - 
                         summary(yfsegg_gams_t[[1]][[3]])$scale) / summary(yfs_egg_base_t)$scale
var_ratio_yfsegg_space_t #-0.0282
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
yfsegg_grids_t <- threepanel_base_mse(yfs_egg, yfs_egg_base_t)
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(yfs_egg, yfsegg_grids_t, ratio_yfsegg_geography_t,
                 var_ratio_yfsegg_space_t,
                 ratio_yfsegg_phenology_t,
                 var_ratio_yfsegg_month_t)
dev.copy(jpeg,
         here('results/yellowfin_hindcast', 
              'yellowfin_egg_MSE_t.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
```{r}
yfsegg_vcgrids_t <- vc_grids_presentation(yfs_egg, yfsegg_gams_t)
```

Create three panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc_presentation(yfs_egg, yfsegg_gams_t[[1]][[2]], yfsegg_vcgrids_t, "Spatial GAM hindcast")
legend(40,
       2,
       legend = c('Mean SST', '+ 1C'),
       col = c('#000000', '#E69F00'),
       lty = 1,
       bty = 'n',
       lwd = 2,
       cex = 1.2)
dev.copy(jpeg,
         here('results/yellowfin_hindcast',
         'yellowfin_egg_vc_t.jpg'),
         height = 3.5,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

```{r}
tiff(here('results/yellowfin_hindcast/outputs',
         'yellowfin_egg_output.jpg'),
     units = "in",
     width = 40,
     height = 12,
     res = 200)

par(mfrow = c(1, 3),
    mar = c(11, 15, .5, 0.6) + 0.1,
    oma = c(3, 1, 1, 1),
    mgp = c(9, 4, 0))
plot_variable(yfsegg_gams_t,
              covariate = 1,
              bounds = c(-20, 10),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(yfsegg_gams_t,
              covariate = 3,
              bounds = c(-20, 10),
              "Temperature",
              " ",
              "n")
plot_variable(yfsegg_gams_t,
              covariate = 4,
              bounds = c(-20, 10),
              "Salinity",
              " ",
              "n")

dev.off()
```

#### Larvae

Models with flexible phenology or geography 
Note variable coefficient by year for each
```{r}
yfs_larvae_phenology_t <- gam(catch ~ factor(year) +
                                s(doy) +
                                s(lon, lat) +
                                s(doy, by = factor(year)),
                              data = yfs_larvae,
                              family = tw(link = 'log'),
                              method =  'REML')
summary(yfs_larvae_phenology_t)
AIC(yfs_larvae_phenology_t)
# AIC: 8269

yfs_larvae_geography_t <- gam(catch ~ factor(year) +
                                s(lon, lat) +
                                s(lon, lat, by = factor(year)) +
                                s(doy),
                              data = yfs_larvae,
                              family = tw(link = 'log'),
                              method = 'REML')
summary(yfs_larvae_geography_t)
AIC(yfs_larvae_geography_t)
# AIC: 7970
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_yfslarvae_phenology_t <- (summary(yfs_larvae_base_t)$scale - 
                      summary(yfs_larvae_phenology_t)$scale) / summary(yfs_larvae_base_t)$scale
ratio_yfslarvae_phenology_t # 0.166

ratio_yfslarvae_geography_t <- (summary(yfs_larvae_base_t)$scale - 
                      summary(yfs_larvae_geography_t)$scale) / summary(yfs_larvae_base_t)$scale
ratio_yfslarvae_geography_t # 0.371
```

Add ROMS mean temperatures Use a representative area with averaged values from Feb - April for each year This can then be matched to the larvae/larvae data for use in variable coefficient by SST instead of year as above

Determine best GAM from 7 formulations (see function)
```{r}
yfslarvae_gams_t <- gam_selection(yfs_larvae)
AIC(yfslarvae_gams_t[[1]][[1]]) # 7932
AIC(yfslarvae_gams_t[[1]][[2]]) # 7891
AIC(yfslarvae_gams_t[[1]][[3]]) # 7852
summary(yfslarvae_gams_t[[2]]) # best is spatial VC, keep all smooth terms
AIC(yfslarvae_gams_t[[1]][[1]]) - AIC(yfslarvae_gams_t[[2]]) # 80.3
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(yfslarvae_gams_t[[1]][[2]], select = 1, main = "DOY")
plot(yfslarvae_gams_t[[1]][[2]], select = 3, main = "Temperature")
plot(yfslarvae_gams_t[[1]][[2]], select = 4, main = "Salinity")
plot(yfslarvae_gams_t[[1]][[2]], select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/yellowfin_hindcast', 'yellowfin_larvae_roms_monthvc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(yfslarvae_gams_t[[1]][[2]])
dev.copy(jpeg, here('results/yellowfin_hindcast', 'yellowfin_larvae_roms_monthvc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(yfslarvae_gams_t[[1]][[3]], select = 1, main = "DOY")
plot(yfslarvae_gams_t[[1]][[3]], select = 3, main = "Temperature")
plot(yfslarvae_gams_t[[1]][[3]], select = 4, main = "Salinity")
plot(yfslarvae_gams_t[[1]][[3]], select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/yellowfin_hindcast', 'yellowfin_larvae_roms_spacevc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(yfslarvae_gams_t[[1]][[3]])
dev.copy(jpeg, here('results/yellowfin_hindcast', 'yellowfin_larvae_roms_spacevc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_yfslarvae_month_t <- (summary(yfs_larvae_base_t)$scale - 
                         summary(yfslarvae_gams_t[[1]][[2]])$scale) / summary(yfs_larvae_base_t)$scale
var_ratio_yfslarvae_month_t #0.078

var_ratio_yfslarvae_space_t <- (summary(yfs_larvae_base_t)$scale - 
                         summary(yfslarvae_gams_t[[1]][[3]])$scale) / summary(yfs_larvae_base_t)$scale
var_ratio_yfslarvae_space_t #0.956
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
yfslarvae_grids_t <- threepanel_base_mse(yfs_larvae, yfs_larvae_base_t)
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(yfs_larvae, yfslarvae_grids_t, ratio_yfslarvae_geography_t,
                 var_ratio_yfslarvae_space_t,
                 ratio_yfslarvae_phenology_t,
                 var_ratio_yfslarvae_month_t)
dev.copy(jpeg,
         here('results/yellowfin_hindcast', 
              'yellowfin_larvae_MSE_t.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
```{r}
yfslarvae_vcgrids_t <- vc_grids_presentation(yfs_larvae, yfslarvae_gams_t)
```

Create three panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc_presentation(yfs_larvae, yfslarvae_gams_t[[2]], yfslarvae_vcgrids_t, "Spatial GAM hindcast")
legend(140,
       21,
       legend = c('Mean SST', '+ 1C'),
       col = c('#000000', '#E69F00'),
       lty = 1,
       bty = 'n',
       lwd = 2,
       cex = 1.2)
dev.copy(jpeg,
         here('results/yellowfin_hindcast',
         'yellowfin_larvae_vc_t.jpg'),
         height = 3.5,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

```{r}
tiff(here('results/yellowfin_hindcast/outputs',
         'yellowfin_larvae_output.jpg'),
     units = "in",
     width = 40,
     height = 12,
     res = 200)

par(mfrow = c(1, 3),
    mar = c(11, 15, .5, 0.6) + 0.1,
    oma = c(3, 1, 1, 1),
    mgp = c(9, 4, 0))
plot_variable(yfslarvae_gams_t,
              covariate = 1,
              bounds = c(-20, 10),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(yfslarvae_gams_t,
              covariate = 3,
              bounds = c(-20, 10),
              "Temperature",
              " ",
              "n")
plot_variable(yfslarvae_gams_t,
              covariate = 4,
              bounds = c(-20, 10),
              "Salinity",
              " ",
              "n")

dev.off()
```

## Alaska Plaice
### Tweedie
```{r}
akp_egg_base_t <- gam(catch ~ factor(year) +
                        s(lon, lat) +
                        s(doy),
                      data = akp_egg,
                      family = tw(link = 'log'),
                      method =  'REML')
summary(akp_egg_base_t)

akp_larvae_base_t <- gam(catch ~ factor(year) +
                           s(lon, lat) +
                           s(doy),
                         data = akp_larvae,
                         family = tw(link = 'log'),
                         method = 'REML')
summary(akp_larvae_base_t)
```

#### Eggs

Models with flexible phenology or geography 
Note variable coefficient by year for each
This takes forever to run, don't do unless necessary to recreate figure
```{r}
akp_egg_phenology_t <- gam(catch ~ factor(year) +
                             s(doy) +
                             s(lon, lat) +
                             s(doy, by = factor(year)),
                           data = akp_egg,
                           family = tw(link = 'log'),
                           method =  'REML')
summary(akp_egg_phenology_t)
AIC(akp_egg_phenology_t)
# AIC: 13892

akp_egg_geography_t <- gam(catch ~ factor(year) +
                             s(lon, lat) +
                             s(lon, lat, by = factor(year)) +
                             s(doy),
                           data = akp_egg,
                           family = tw(link = 'log'),
                           method =  'REML')
summary(akp_egg_geography_t)
AIC(akp_egg_geography_t)
# AIC: 13005
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_akpegg_phenology_t <- (summary(akp_egg_base_t)$scale - 
                      summary(akp_egg_phenology_t)$scale) / summary(akp_egg_base_t)$scale
ratio_akpegg_phenology_t # 0.195

ratio_akpegg_geography_t <- (summary(akp_egg_base_t)$scale - 
                      summary(akp_egg_geography_t)$scale) / summary(akp_egg_base_t)$scale
ratio_akpegg_geography_t # 0.443
```

Add ROMS mean temperatures Use a representative area with averaged values from Feb - April for each year This can then be matched to the egg/larvae data for use in variable coefficient by SST instead of year as above


Determine best GAM from 7 formulations (see function)
```{r}
akpegg_gams_t <- gam_selection(akp_egg)
AIC(akpegg_gams_t[[1]][[1]]) # 14191
AIC(akpegg_gams_t[[1]][[2]]) # 14140
AIC(akpegg_gams_t[[1]][[3]]) # 14143
summary(akpegg_gams_t[[2]]) # best model is temporal VC, keep all terms
AIC(akpegg_gams_t[[1]][[1]]) - AIC(akpegg_gams_t[[2]]) #50.4
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(akpegg_gams_t[[1]][[2]], select = 1, main = "DOY")
plot(akpegg_gams_t[[1]][[2]], select = 3, main = "Temperature")
plot(akpegg_gams_t[[1]][[2]], select = 4, main = "Salinity")
plot(akpegg_gams_t[[1]][[2]], select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/plaice_hindcast', 'plaice_egg_roms_monthvc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(akpegg_gams_t[[1]][[2]])
dev.copy(jpeg, here('results/plaice_hindcast', 'plaice_egg_roms_monthvc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(akpegg_gams_t[[1]][[3]], select = 1, main = "DOY")
plot(akpegg_gams_t[[1]][[3]], select = 3, main = "Temperature")
plot(akpegg_gams_t[[1]][[3]], select = 4, main = "Salinity")
plot(akpegg_gams_t[[1]][[3]], select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/plaice_hindcast', 'plaice_egg_roms_spacevc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(akpegg_gams_t[[1]][[3]])
dev.copy(jpeg, here('results/plaice_hindcast', 'plaice_egg_roms_spacevc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_akpegg_month_t <- (summary(akp_egg_base_t)$scale - 
                         summary(akpegg_gams_t[[1]][[2]])$scale) / summary(akp_egg_base_t)$scale
var_ratio_akpegg_month_t # 0.122

var_ratio_akpegg_space_t <- (summary(akp_egg_base_t)$scale - 
                         summary(akpegg_gams_t[[1]][[3]])$scale) / summary(akp_egg_base_t)$scale
var_ratio_akpegg_space_t # 0.111
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
akpegg_grids_t <- threepanel_base_mse(akp_egg, akp_egg_base_t)
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(akp_egg, akpegg_grids_t, ratio_akpegg_geography_t,
                 var_ratio_akpegg_space_t,
                 ratio_akpegg_phenology_t,
                 var_ratio_akpegg_month_t)
dev.copy(jpeg,
         here('results/plaice_hindcast', 
              'plaice_egg_MSE_t.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
```{r}
akpegg_vcgrids_t <- vc_grids_presentation(akp_egg, akpegg_gams_t)
```

Create three panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc_presentation(akp_egg, akpegg_gams_t[[2]], akpegg_vcgrids_t, "Spatial GAM hindcast")
legend(120,
       5,
       legend = c('Mean SST', '+ 1C'),
       col = c('#000000', '#E69F00'),
       lty = 1,
       bty = 'n',
       lwd = 2,
       cex = 1.2)
dev.copy(jpeg,
         here('results/plaice_hindcast',
         'plaice_egg_vc_t.jpg'),
         height = 3.5,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

```{r}
tiff(here('results/plaice_hindcast/outputs',
         'plaice_egg_output.jpg'),
     units = "in",
     width = 52,
     height = 12,
     res = 200)

par(mfrow = c(1, 4),
    mar = c(11, 15, .5, 0.6) + 0.1,
    oma = c(3, 1, 1, 1),
    mgp = c(9, 4, 0))
plot_variable(akpegg_gams_t,
              covariate = 1,
              bounds = c(-7, 8),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(akpegg_gams_t,
              covariate = 3,
              bounds = c(-7, 8),
              "Temperature",
              " ",
              "n")
plot_variable(akpegg_gams_t,
              covariate = 4,
              bounds = c(-7, 8),
              "Salinity",
              " ",
              "n")
plot_variable(akpegg_gams_t,
              covariate = 5,
              bounds = c(-7, 8),
              "Day of Year",
              " ",
              "n")
dev.off()
```

#### Larvae

Models with flexible phenology or geography 
Note variable coefficient by year for each
```{r}
akp_larvae_phenology_t <- gam(catch ~ factor(year) +
                                s(doy) +
                                s(lon, lat) +
                                s(doy, by = factor(year)),
                              data = akp_larvae,
                              family = tw(link = 'log'),
                              method = 'REML')
summary(akp_larvae_phenology_t)
AIC(akp_larvae_phenology_t)
# AIC: 6035

akp_larvae_geography_t <- gam(catch ~ factor(year) +
                                s(lon, lat) +
                                s(lon, lat, by = factor(year)) +
                                s(doy),
                              data = akp_larvae,
                              family = tw(link = 'log'),
                              method = 'REML')
summary(akp_larvae_geography_t)
AIC(akp_larvae_geography_t)
# AIC: 5071
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_akplarvae_phenology_t <- (summary(akp_larvae_base_t)$scale - 
                      summary(akp_larvae_phenology_t)$scale) / summary(akp_larvae_base_t)$scale
ratio_akplarvae_phenology_t # 0.132

ratio_akplarvae_geography_t <- (summary(akp_larvae_base_t)$scale - 
                      summary(akp_larvae_geography_t)$scale) / summary(akp_larvae_base_t)$scale
ratio_akplarvae_geography_t # 0.554
```

Add ROMS mean temperatures Use a representative area with averaged values from Feb - April for each year This can then be matched to the larvae/larvae data for use in variable coefficient by SST instead of year as above

Determine best GAM from 7 formulations (see function)
```{r}
akplarvae_gams_t <- gam_selection(akp_larvae)
AIC(akplarvae_gams_t[[1]][[1]]) # 5768
AIC(akplarvae_gams_t[[1]][[2]]) # 5654
AIC(akplarvae_gams_t[[1]][[3]]) # 5660
summary(akplarvae_gams_t[[2]]) # best is phenology VC
AIC(akplarvae_gams_t[[1]][[1]]) - AIC(akplarvae_gams_t[[2]]) #114.1
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(akplarvae_gams_t[[1]][[2]], select = 1, main = "DOY")
plot(akplarvae_gams_t[[1]][[2]], select = 3, main = "Temperature")
plot(akplarvae_gams_t[[1]][[2]], select = 4, main = "Salinity")
plot(akplarvae_gams_t[[1]][[2]], select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/plaice_hindcast', 'plaice_larvae_roms_monthvc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(akplarvae_gams_t[[1]][[2]])
dev.copy(jpeg, here('results/plaice_hindcast', 'plaice_larvae_roms_monthvc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(akplarvae_gams_t[[1]][[3]], select = 1, main = "DOY")
plot(akplarvae_gams_t[[1]][[3]], select = 3, main = "Temperature")
plot(akplarvae_gams_t[[1]][[3]], select = 4, main = "Salinity")
plot(akplarvae_gams_t[[1]][[3]], select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/plaice_hindcast', 'plaice_larvae_roms_spacevc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(akplarvae_gams_t[[1]][[3]])
dev.copy(jpeg, here('results/plaice_hindcast', 'plaice_larvae_roms_spacevc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_akplarvae_month_t <- (summary(akp_larvae_base_t)$scale - 
                         summary(akplarvae_gams_t[[1]][[2]])$scale) / summary(akp_larvae_base_t)$scale
var_ratio_akplarvae_month_t # 0.169

var_ratio_akplarvae_space_t <- (summary(akp_larvae_base_t)$scale - 
                         summary(akplarvae_gams_t[[1]][[3]])$scale) / summary(akp_larvae_base_t)$scale
var_ratio_akplarvae_space_t # 0.166
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
akplarvae_grids_t <- threepanel_base_mse(akp_larvae, akp_larvae_base_t)
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(akp_larvae, akplarvae_grids_t, ratio_akplarvae_geography_t,
                 var_ratio_akplarvae_space_t,
                 ratio_akplarvae_phenology_t,
                 var_ratio_akplarvae_month_t)
dev.copy(jpeg,
         here('results/plaice_hindcast', 
              'plaice_larvae_MSE_t.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
```{r}
akplarvae_vcgrids_t <- vc_grids_presentation(akp_larvae, akplarvae_gams_t)
```

Create three panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc_presentation(akp_larvae, akplarvae_gams_t[[2]], akplarvae_vcgrids_t, "Spatial GAM hindcast")
legend(140,
       8,
       legend = c('Mean SST', '+ 1C'),
       col = c('#000000', '#E69F00'),
       lty = 1,
       bty = 'n',
       lwd = 2,
       cex = 1.2)
dev.copy(jpeg,
         here('results/plaice_hindcast',
         'plaice_larvae_vc_t.jpg'),
         height = 3.5,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

```{r}
tiff(here('results/plaice_hindcast/outputs',
         'plaice_larvae_output.jpg'),
     units = "in",
     width = 52,
     height = 12,
     res = 200)

par(mfrow = c(1, 4),
    mar = c(11, 15, .5, 0.6) + 0.1,
    oma = c(3, 1, 1, 1),
    mgp = c(9, 4, 0))
plot_variable(akplarvae_gams_t,
              covariate = 1,
              bounds = c(-7, 8),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(akplarvae_gams_t,
              covariate = 3,
              bounds = c(-7, 8),
              "Temperature",
              " ",
              "n")
plot_variable(akplarvae_gams_t,
              covariate = 4,
              bounds = c(-7, 8),
              "Salinity",
              " ",
              "n")
plot_variable(akplarvae_gams_t,
              covariate = 5,
              bounds = c(-7, 8),
              "Day of Year",
              " ",
              "n")
dev.off()
```


## Pacific Cod
### Tweedie
```{r}
pcod_larvae_base_t <- gam(catch ~ factor(year) +
                           s(lon, lat) +
                           s(doy),
                         data = pcod_larvae,
                         family = tw(link = 'log'),
                         method = 'REML')
summary(pcod_larvae_base_t)
```

#### Larvae

Models with flexible phenology or geography 
Note variable coefficient by year for each
```{r}
pcod_larvae_phenology_t <- gam(catch ~ factor(year) +
                                s(doy) +
                                s(lon, lat) +
                                s(doy, by = factor(year)),
                              data = pcod_larvae,
                              family = tw(link = 'log'),
                              method = 'REML')
summary(pcod_larvae_phenology_t)
AIC(pcod_larvae_phenology_t)
# AIC: 8606

pcod_larvae_geography_t <- gam(catch ~ factor(year) +
                                s(lon, lat) +
                                s(lon, lat, by = factor(year)) +
                                s(doy),
                              data = pcod_larvae,
                              family = tw(link = 'log'),
                              method = 'REML')
summary(pcod_larvae_geography_t)
AIC(pcod_larvae_geography_t)
# AIC: 8237
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_pcodlarvae_phenology_t <- (summary(pcod_larvae_base_t)$scale - 
                      summary(pcod_larvae_phenology_t)$scale) / summary(pcod_larvae_base_t)$scale
ratio_pcodlarvae_phenology_t # 0.148

ratio_pcodlarvae_geography_t <- (summary(pcod_larvae_base_t)$scale - 
                      summary(pcod_larvae_geography_t)$scale) / summary(pcod_larvae_base_t)$scale
ratio_pcodlarvae_geography_t # 0.313
```

Add ROMS mean temperatures Use a representative area with averaged values from Feb - April for each year This can then be matched to the larvae/larvae data for use in variable coefficient by SST instead of year as above

Determine best GAM from 7 formulations (see function)
```{r}
pcodlarvae_gams_t <- gam_selection(pcod_larvae)
AIC(pcodlarvae_gams_t[[1]][[1]]) # 8862
AIC(pcodlarvae_gams_t[[1]][[2]]) # 8808
AIC(pcodlarvae_gams_t[[1]][[3]]) # 8808
summary(pcodlarvae_gams_t[[2]]) # best is phenology VC
AIC(pcodlarvae_gams_t[[1]][[1]]) - AIC(pcodlarvae_gams_t[[2]]) # 54
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(pcodlarvae_gams_t[[1]][[2]], select = 1, main = "DOY")
plot(pcodlarvae_gams_t[[1]][[2]], select = 3, main = "Temperature")
plot(pcodlarvae_gams_t[[1]][[2]], select = 4, main = "Salinity")
plot(pcodlarvae_gams_t[[1]][[2]], select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/cod_hindcast', 'cod_larvae_roms_monthvc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(pcodlarvae_gams_t[[1]][[2]])
dev.copy(jpeg, here('results/cod_hindcast', 'cod_larvae_roms_monthvc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(pcodlarvae_gams_t[[1]][[3]], select = 1, main = "DOY")
plot(pcodlarvae_gams_t[[1]][[3]], select = 3, main = "Temperature")
plot(pcodlarvae_gams_t[[1]][[3]], select = 4, main = "Salinity")
plot(pcodlarvae_gams_t[[1]][[3]], select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/cod_hindcast', 'cod_larvae_roms_spacevc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(pcodlarvae_gams_t[[1]][[3]])
dev.copy(jpeg, here('results/cod_hindcast', 'cod_larvae_roms_spacevc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_pcodlarvae_month_t <- (summary(pcod_larvae_base_t)$scale - 
                         summary(pcodlarvae_gams_t[[1]][[2]])$scale) / summary(pcod_larvae_base_t)$scale
var_ratio_pcodlarvae_month_t # 0.06

var_ratio_pcodlarvae_space_t <- (summary(pcod_larvae_base_t)$scale - 
                         summary(pcodlarvae_gams_t[[1]][[3]])$scale) / summary(pcod_larvae_base_t)$scale
var_ratio_pcodlarvae_space_t #  0.06
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
pcodlarvae_grids_t <- threepanel_base_mse(pcod_larvae, pcod_larvae_base_t)
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(pcod_larvae, pcodlarvae_grids_t, ratio_pcodlarvae_geography_t,
                 var_ratio_pcodlarvae_space_t,
                 ratio_pcodlarvae_phenology_t,
                 var_ratio_pcodlarvae_month_t)
dev.copy(jpeg,
         here('results/cod_hindcast', 
              'cod_larvae_MSE_t.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
```{r}
pcodlarvae_vcgrids_t <- vc_grids_presentation(pcod_larvae, pcodlarvae_gams_t)
```

Create three panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc_presentation(pcod_larvae, pcodlarvae_gams_t[[2]], pcodlarvae_vcgrids_t, "Spatial GAM hindcast")
legend(100,
       -1,
       legend = c('Mean SST', '+ 1C'),
       col = c('#000000', '#E69F00'),
       lty = 1,
       bty = 'n',
       lwd = 2,
       cex = 1.2)
dev.copy(jpeg,
         here('results/cod_hindcast',
         'cod_larvae_vc_t.jpg'),
         height = 3.5,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

```{r}
tiff(here('results/cod_hindcast/outputs',
         'cod_larvae_output.jpg'),
     units = "in",
     width = 52,
     height = 12,
     res = 200)

par(mfrow = c(1, 4),
    mar = c(11, 15, .5, 0.6) + 0.1,
    oma = c(3, 1, 1, 1),
    mgp = c(9, 4, 0))
plot_variable(pcodlarvae_gams_t,
              covariate = 1,
              bounds = c(-6, 2.5),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(pcodlarvae_gams_t,
              covariate = 3,
              bounds = c(-6, 2.5),
              "Temperature",
              " ",
              "n")
plot_variable(pcodlarvae_gams_t,
              covariate = 4,
              bounds = c(-6, 2.5),
              "Salinity",
              " ",
              "n")
plot_variable(pcodlarvae_gams_t,
              covariate = 5,
              bounds = c(-6, 2.5),
              "Day of Year",
              " ",
              "n")
dev.off()
```

## Northern Rock Sole
### Tweedie
```{r}
nrs_larvae_base_t <- gam(catch ~ factor(year) +
                           s(lon, lat) +
                           s(doy),
                         data = nrs_larvae,
                         family = tw(link = 'log'),
                         method = 'REML')
summary(nrs_larvae_base_t)
```

#### Larvae

Models with flexible phenology or geography 
Note variable coefficient by year for each
```{r}
nrs_larvae_phenology_t <- gam(catch ~ factor(year) +
                                s(doy) +
                                s(lon, lat) +
                                s(doy, by = factor(year)),
                              data = nrs_larvae,
                              family = tw(link = 'log'),
                              method = 'REML')
summary(nrs_larvae_phenology_t)
AIC(nrs_larvae_phenology_t)
# AIC: 15687

nrs_larvae_geography_t <- gam(catch ~ factor(year) +
                                s(lon, lat) +
                                s(lon, lat, by = factor(year)) +
                                s(doy),
                              data = nrs_larvae,
                              family = tw(link = 'log'),
                              method = 'REML')
summary(nrs_larvae_geography_t)
AIC(nrs_larvae_geography_t)
# AIC: 14868
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_nrslarvae_phenology_t <- (summary(nrs_larvae_base_t)$scale - 
                      summary(nrs_larvae_phenology_t)$scale) / summary(nrs_larvae_base_t)$scale
ratio_nrslarvae_phenology_t # 0.073

ratio_nrslarvae_geography_t <- (summary(nrs_larvae_base_t)$scale - 
                      summary(nrs_larvae_geography_t)$scale) / summary(nrs_larvae_base_t)$scale
ratio_nrslarvae_geography_t # 0.334
```

Add ROMS mean temperatures Use a representative area with averaged values from Feb - April for each year This can then be matched to the larvae/larvae data for use in variable coefficient by SST instead of year as above

Determine best GAM from 7 formulations (see function)
```{r}
nrslarvae_gams_t <- gam_selection(nrs_larvae)
AIC(nrslarvae_gams_t[[1]][[1]]) # 15633
AIC(nrslarvae_gams_t[[1]][[2]]) # 15623
AIC(nrslarvae_gams_t[[1]][[3]]) # 15589
summary(nrslarvae_gams_t[[2]]) # best is location VC
AIC(nrslarvae_gams_t[[1]][[1]]) - AIC(nrslarvae_gams_t[[2]]) # 43.4
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(nrslarvae_gams_t[[1]][[2]], select = 1, main = "DOY")
plot(nrslarvae_gams_t[[1]][[2]], select = 3, main = "Temperature")
plot(nrslarvae_gams_t[[1]][[2]], select = 4, main = "Salinity")
plot(nrslarvae_gams_t[[1]][[2]], select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/rocksole_hindcast', 'rocksole_larvae_roms_monthvc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(nrslarvae_gams_t[[1]][[2]])
dev.copy(jpeg, here('results/rocksole_hindcast', 'rocksole_larvae_roms_monthvc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(nrslarvae_gams_t[[1]][[3]], select = 1, main = "DOY")
plot(nrslarvae_gams_t[[1]][[3]], select = 3, main = "Temperature")
plot(nrslarvae_gams_t[[1]][[3]], select = 4, main = "Salinity")
plot(nrslarvae_gams_t[[1]][[3]], select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/rocksole_hindcast', 'rocksole_larvae_roms_spacevc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(nrslarvae_gams_t[[1]][[3]])
dev.copy(jpeg, here('results/rocksole_hindcast', 'rocksole_larvae_roms_spacevc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_nrslarvae_month_t <- (summary(nrs_larvae_base_t)$scale - 
                         summary(nrslarvae_gams_t[[1]][[2]])$scale) / summary(nrs_larvae_base_t)$scale
var_ratio_nrslarvae_month_t # 0.079

var_ratio_nrslarvae_space_t <- (summary(nrs_larvae_base_t)$scale - 
                         summary(nrslarvae_gams_t[[1]][[3]])$scale) / summary(nrs_larvae_base_t)$scale
var_ratio_nrslarvae_space_t # 0.089
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
nrslarvae_grids_t <- threepanel_base_mse(nrs_larvae, nrs_larvae_base_t)
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(nrs_larvae, nrslarvae_grids_t, ratio_nrslarvae_geography_t,
                 var_ratio_nrslarvae_space_t,
                 ratio_nrslarvae_phenology_t,
                 var_ratio_nrslarvae_month_t)
dev.copy(jpeg,
         here('results/rocksole_hindcast', 
              'rocksole_larvae_MSE_t.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
```{r}
nrslarvae_vcgrids_t <- vc_grids_presentation(nrs_larvae, nrslarvae_gams_t)
```

Create three panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc_presentation(nrs_larvae, nrslarvae_gams_t[[2]], nrslarvae_vcgrids_t, "Spatial GAM hindcast")
legend(110,
       1,
       legend = c('Mean SST', '+ 1C'),
       col = c('#000000', '#E69F00'),
       lty = 1,
       bty = 'n',
       lwd = 2,
       cex = 1.2)
dev.copy(jpeg,
         here('results/rocksole_hindcast',
         'rocksole_larvae_vc_t.jpg'),
         height = 3.5,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

```{r}
tiff(here('results/rocksole_hindcast/outputs',
         'rocksole_larvae_output.jpg'),
     units = "in",
     width = 40,
     height = 12,
     res = 200)

par(mfrow = c(1, 3),
    mar = c(11, 15, .5, 0.6) + 0.1,
    oma = c(3, 1, 1, 1),
    mgp = c(9, 4, 0))
plot_variable(nrslarvae_gams_t,
              covariate = 1,
              bounds = c(-2.5, 1),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(nrslarvae_gams_t,
              covariate = 3,
              bounds = c(-2.5, 1),
              "Temperature",
              " ",
              "n")
plot_variable(nrslarvae_gams_t,
              covariate = 4,
              bounds = c(-2.5, 1),
              "Salinity",
              " ",
              "n")
dev.off()
```


## GAM multipanel figures
```{r}
# Figure containing location or doy by mean sst
tiff(here('results',
          'hindcast_output.jpg'),
     units = "in",
     width = 65,
     height = 90,
     res = 300)
par(mfrow = c(7, 5),
    mar = c(11, 15, 5, 1.5) + 0.1,
    oma = c(3, 35, 15, 1),
    mgp = c(10, 4, 0),
    family = "serif")
plot_variable(pkegg_gams_t,
              covariate = 1,
              bounds = c(-5.5, 7),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(pkegg_gams_t,
              covariate = 3,
              bounds = c(-5.5, 7),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(pkegg_gams_t,
              covariate = 4,
              bounds = c(-5.5, 7),
              "Salinity",
              " ",
              "n")
plot(0, xaxt = "n", yaxt = "n", bty = "n", pch = "", ylab = "", xlab = "")
location_plot(pkegg_gams_t,
              pk_egg)
plot_variable(pklarvae_gams_t,
              covariate = 1,
              bounds = c(-5.5, 7),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(pklarvae_gams_t,
              covariate = 3,
              bounds = c(-5.5, 7),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(pklarvae_gams_t,
              covariate = 4,
              bounds = c(-5.5, 7),
              "Salinity",
              " ",
              "n")
plot(0, xaxt = "n", yaxt = "n", bty = "n", pch = "", ylab = "", xlab = "")
location_plot(pklarvae_gams_t,
              pk_larvae)
plot_variable(fhsegg_gams_t,
              covariate = 1,
              bounds = c(-6, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(fhsegg_gams_t,
              covariate = 3,
              bounds = c(-6, 3),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(fhsegg_gams_t,
              covariate = 4,
              bounds = c(-6, 3),
              "Salinity",
              " ",
              "n")
plot_variable(fhsegg_gams_t,
              covariate = 5,
              bounds = c(-6, 3),
              "Day of Year",
              " ",
              "n")
plot(0, xaxt = "n", yaxt = "n", bty = "n", pch = "", ylab = "", xlab = "")
plot_variable(fhslarvae_gams_t,
              covariate = 1,
              bounds = c(-6, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(fhslarvae_gams_t,
              covariate = 3,
              bounds = c(-6, 3),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(fhslarvae_gams_t,
              covariate = 4,
              bounds = c(-6, 3),
              "Salinity",
              " ",
              "n")
plot(0, xaxt = "n", yaxt = "n", bty = "n", pch = "", ylab = "", xlab = "")
location_plot(fhslarvae_gams_t,
              fhs_larvae)
plot_variable(akpegg_gams_t,
              covariate = 1,
              bounds = c(-7, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(akpegg_gams_t,
              covariate = 3,
              bounds = c(-7, 3),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(akpegg_gams_t,
              covariate = 4,
              bounds = c(-7, 3),
              "Salinity",
              " ",
              "n")
plot_variable(akpegg_gams_t,
              covariate = 5,
              bounds = c(-7, 3),
              "Day of Year",
              " ",
              "n")
plot(0, xaxt = "n", yaxt = "n", bty = "n", pch = "", ylab = "", xlab = "")
plot_variable(akplarvae_gams_t,
              covariate = 1,
              bounds = c(-7, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(akplarvae_gams_t,
              covariate = 3,
              bounds = c(-7, 3),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(akplarvae_gams_t,
              covariate = 4,
              bounds = c(-7, 3),
              "Salinity",
              " ",
              "n")
plot_variable(akplarvae_gams_t,
              covariate = 5,
              bounds = c(-7, 3),
              "Day of Year",
              " ",
              "n")
plot(0, xaxt = "n", yaxt = "n", bty = "n", pch = "", ylab = "", xlab = "")
plot_variable(yfslarvae_gams_t,
              covariate = 1,
              bounds = c(-4, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(yfslarvae_gams_t,
              covariate = 3,
              bounds = c(-4, 3),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(yfslarvae_gams_t,
              covariate = 4,
              bounds = c(-4, 3),
              "Salinity",
              " ",
              "n")
plot(0, xaxt = "n", yaxt = "n", bty = "n", pch = "", ylab = "", xlab = "")
location_plot(yfslarvae_gams_t,
              yfs_larvae)
mtext("Walleye Pollock", 
      side = 2, 
      line = 25, 
      outer = TRUE, 
      cex = 7,
      at = 0.86)
mtext("Flathead Sole", 
      side = 2, 
      line = 25, 
      outer = TRUE, 
      cex = 7,
      at = 0.58)
mtext("Alaska Plaice", 
      side = 2, 
      line = 25, 
      outer = TRUE, 
      cex = 7,
      at = 0.29)
mtext("Yellowfin Sole", 
      side = 2, 
      line = 25, 
      outer = TRUE, 
      cex = 7,
      at = 0.08)
mtext("Day of Year", 
      side = 3, 
      line = 2, 
      outer = TRUE, 
      cex = 7,
      at = 0.11)
mtext("Temperature", 
      side = 3, 
      line = 2, 
      outer = TRUE, 
      cex = 7,
      at = 0.31)
mtext("Salinity", 
      side = 3, 
      line = 2, 
      outer = TRUE, 
      cex = 7,
      at = 0.51)
mtext("Day of Year \n by Mean SST",
      side = 3,
      line = -1.8,
      outer = TRUE,
      cex = 7,
      at = 0.71)
mtext("Location",
      side = 3,
      line = 2,
      outer = TRUE,
      cex = 7,
      at = 0.91)
mtext("Eggs", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.93)
mtext("Larvae", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.79)
mtext("Eggs", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.65)
mtext("Larvae", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.51)
mtext("Eggs", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.36)
mtext("Larvae", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.22)
mtext("Larvae", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.08)
dev.off()
```


```{r}
# Figure containing 
tiff(here('results',
          'hindcast_output2.jpg'),
     units = "in",
     width = 65,
     height = 90,
     res = 300)
par(mfrow = c(7, 5),
    mar = c(11, 15, 5, 1.5) + 0.1,
    oma = c(3, 35, 15, 1),
    mgp = c(10, 4, 0),
    family = "serif")
plot_variable(pkegg_gams_t,
              covariate = 1,
              bounds = c(-5.5, 7),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(pkegg_gams_t,
              covariate = 3,
              bounds = c(-5.5, 7),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(pkegg_gams_t,
              covariate = 4,
              bounds = c(-5.5, 7),
              "Salinity",
              " ",
              "n")
phenology_change(pkegg_vcgrids_t)
distribution_change(pk_egg,
                    pkegg_vcgrids_t)
plot_variable(pklarvae_gams_t,
              covariate = 1,
              bounds = c(-5.5, 7),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(pklarvae_gams_t,
              covariate = 3,
              bounds = c(-5.5, 7),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(pklarvae_gams_t,
              covariate = 4,
              bounds = c(-5.5, 7),
              "Salinity",
              " ",
              "n")
phenology_change(pklarvae_vcgrids_t)
distribution_change(pk_larvae,
                    pklarvae_vcgrids_t)
plot_variable(fhsegg_gams_t,
              covariate = 1,
              bounds = c(-6, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(fhsegg_gams_t,
              covariate = 3,
              bounds = c(-6, 3),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(fhsegg_gams_t,
              covariate = 4,
              bounds = c(-6, 3),
              "Salinity",
              " ",
              "n")
phenology_change(fhsegg_vcgrids_t)
distribution_change(fhs_egg,
                    fhsegg_vcgrids_t)
plot_variable(fhslarvae_gams_t,
              covariate = 1,
              bounds = c(-6, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(fhslarvae_gams_t,
              covariate = 3,
              bounds = c(-6, 3),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(fhslarvae_gams_t,
              covariate = 4,
              bounds = c(-6, 3),
              "Salinity",
              " ",
              "n")
phenology_change(fhslarvae_vcgrids_t)
distribution_change(fhs_larvae,
                    fhslarvae_vcgrids_t)
plot_variable(akpegg_gams_t,
              covariate = 1,
              bounds = c(-7, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(akpegg_gams_t,
              covariate = 3,
              bounds = c(-7, 3),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(akpegg_gams_t,
              covariate = 4,
              bounds = c(-7, 3),
              "Salinity",
              " ",
              "n")
phenology_change(akpegg_vcgrids_t)
distribution_change(akp_egg,
                    akpegg_vcgrids_t)
plot_variable(akplarvae_gams_t,
              covariate = 1,
              bounds = c(-7, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(akplarvae_gams_t,
              covariate = 3,
              bounds = c(-7, 3),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(akplarvae_gams_t,
              covariate = 4,
              bounds = c(-7, 3),
              "Salinity",
              " ",
              "n")
phenology_change(akplarvae_vcgrids_t)
distribution_change(akp_larvae,
                    akplarvae_vcgrids_t)
plot_variable(yfslarvae_gams_t,
              covariate = 1,
              bounds = c(-4, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(yfslarvae_gams_t,
              covariate = 3,
              bounds = c(-4, 3),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(yfslarvae_gams_t,
              covariate = 4,
              bounds = c(-4, 3),
              "Salinity",
              " ",
              "n")
phenology_change(yfslarvae_vcgrids_t)
distribution_change(yfs_larvae,
                    yfslarvae_vcgrids_t)
mtext("Walleye Pollock", 
      side = 2, 
      line = 25, 
      outer = TRUE, 
      cex = 7,
      at = 0.86)
mtext("Flathead Sole", 
      side = 2, 
      line = 25, 
      outer = TRUE, 
      cex = 7,
      at = 0.58)
mtext("Alaska Plaice", 
      side = 2, 
      line = 25, 
      outer = TRUE, 
      cex = 7,
      at = 0.29)
mtext("Yellowfin Sole", 
      side = 2, 
      line = 25, 
      outer = TRUE, 
      cex = 7,
      at = 0.08)
mtext("Day of Year", 
      side = 3, 
      line = 2, 
      outer = TRUE, 
      cex = 7,
      at = 0.11)
mtext("Temperature", 
      side = 3, 
      line = 2, 
      outer = TRUE, 
      cex = 7,
      at = 0.31)
mtext("Salinity", 
      side = 3, 
      line = 2, 
      outer = TRUE, 
      cex = 7,
      at = 0.51)
mtext("Phenology Change",
      side = 3,
      line = 2,
      outer = TRUE,
      cex = 7,
      at = 0.71)
mtext("Location Change",
      side = 3,
      line = 2,
      outer = TRUE,
      cex = 7,
      at = 0.91)
mtext("Eggs", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.93)
mtext("Larvae", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.79)
mtext("Eggs", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.65)
mtext("Larvae", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.51)
mtext("Eggs", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.36)
mtext("Larvae", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.22)
mtext("Larvae", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.08)
dev.off()
```