---
title: "Hindcast ROMS variables"
author: "Rebecca Howard"
date: "9/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(marmap)
library(maps)
library(mapdata)
library(fields)
library(here)
library(mgcv)
library(tidyr)
library(dplyr)
library(ggplot2)
library(raster)
library(rgdal)
library(viridis)
source(here('code/functions', 'vis_gam_COLORS.R'))
source(here('code/functions', 'distance_function.R'))
```

In-script functions for maps and data
```{r}
# Function to make maps
map_phenology <- function(data, grid, grid2){
  nlat = 80
  nlon = 120
  latd = seq(min(grid$lat), max(grid$lat), length.out = nlat)
  lond = seq(min(grid$lon), max(grid$lon), length.out = nlon)
  my_color = colorRampPalette(c("#1C0D51", "#4C408E", "#7E77B0",
                                "#AFABCB", "#DAD9E5", "#F9F9F9",
                                "#FFDAB7", "#FFB377","#E18811",
                                "#AC6000", "#743700"))
  color_levels = 100
  max_absolute_value = max(abs(c(min(grid$pred, na.rm = T), max(grid$pred, na.rm = T)))) 
  color_sequence = seq(-max_absolute_value, max_absolute_value, 
                       length.out = color_levels + 1)
  n_in_class = hist(grid$pred, breaks = color_sequence, plot = F)$counts > 0
  col_to_include = min(which(n_in_class == T)):max(which(n_in_class == T))
  breaks_to_include = min(which(n_in_class == T)):(max(which(n_in_class == T)) + 1)
  image(lond,
        latd,
        t(matrix(grid$pred,
                 nrow = length(latd),
                 ncol = length(lond),
                 byrow = T)),
        xlim = c(-176.5, -156.5),
        ylim = c(52, 62),
        axes = FALSE,
        xlab = "",
        ylab = "")
  rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "mintcream")
  par(new = TRUE)
  image(lond,
        latd,
        t(matrix(grid$pred,
                 nrow = length(latd),
                 ncol = length(lond),
                 byrow = T)),
        col = my_color(n = color_levels)[col_to_include], 
        breaks = color_sequence[breaks_to_include],
        ylab = "Latitude",
        xlab = "Longitude",
        xlim = c(-176.5, -156.5),
        ylim = c(52, 62),
        main = "Distribution",
        cex.main = 1.2,
        cex.lab = 1.1,
        cex.axis = 1.1)
  symbols(data$lon[data$larvalcatchper10m2 > 0],
          data$lat[data$larvalcatchper10m2 > 0],
          circles = log(data$larvalcatchper10m2 + 1)[data$larvalcatchper10m2 > 0],
          inches = 0.1,
          bg = alpha('grey', 0.1),
          fg = alpha('black', 0.05),
          add = T)
  points(data$lon[data$larvalcatchper10m2 == 0], 
         data$lat[data$larvalcatchper10m2 == 0], 
         pch = '')
  maps::map("worldHires",
            fill = T,
            col = "wheat4",
            add = T)
  image.plot(legend.only = T,
             col = my_color(n = color_levels)[col_to_include],
             legend.shrink = 0.2,
             smallplot = c(.765, .79, .25, .37),
             legend.cex = 0.4,
             axis.args = list(cex.axis = 0.6),
             legend.width = 0.5,
             legend.mar = 6,
             zlim = c(min(grid$pred, na.rm = T), max(grid$pred, na.rm = T)),
             legend.args = list("Predicted \n Change",
                                side = 2, cex = 0.7))
  plot(grid2$doy,
       grid2$pred,
       main = 'Phenology',
       type = 'l',
       ylab = 'Egg density ln(n/10m2)',
       xlab = 'Day of the year',
       cex.lab = 1.1,
       cex.axis = 1.1,
       cex.main = 1.2,
       xlim = c(min(grid2$doy, na.rm = T), max(grid2$doy, na.rm = T)),
       ylim = range(c(grid2$pred_up, grid2$pred_lw)),
       col = 'blue',
       lwd = 2)
  polygon(c(grid2$doy, rev(grid2$doy)),
          c(grid2$pred_lw, rev(grid2$pred_up)),
          col = alpha('gray', 0.6),
          lty = 0)
}
map_vc <- function(data, grid, grid2){
  nlat = 80
  nlon = 120
  latd = seq(min(pk_egg$lat), max(pk_egg$lat), length.out = nlat)
  lond = seq(min(pk_egg$lon), max(pk_egg$lon), length.out = nlon)
  my_color = colorRampPalette(c("#1C0D51", "#4C408E", "#7E77B0",
                                "#AFABCB", "#DAD9E5", "#F9F9F9",
                                "#FFDAB7", "#FFB377","#E18811",
                                "#AC6000", "#743700"))
  color_levels = 100
  max_absolute_value = max(abs(c(min(grid$diff, na.rm = T), max(grid$diff, na.rm = T)))) 
  color_sequence = seq(-max_absolute_value, max_absolute_value, 
                       length.out = color_levels + 1)
  n_in_class = hist(grid$diff, breaks = color_sequence, plot = F)$counts > 0
  col_to_include = min(which(n_in_class == T)):max(which(n_in_class == T))
  breaks_to_include = min(which(n_in_class == T)):(max(which(n_in_class == T)) + 1)
  image(lond,
        latd,
        t(matrix(grid$diff,
                 nrow = length(latd),
                 ncol = length(lond),
                 byrow = T)),
        xlim = c(-176.5, -156.5),
        ylim = c(52, 62),
        axes = FALSE,
        xlab = "",
        ylab = "")
  rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "mintcream")
  par(new = TRUE)
  image(lond,
        latd,
        t(matrix(grid$diff,
                 nrow = length(latd),
                 ncol = length(lond),
                 byrow = T)),
        col = my_color(n = color_levels)[col_to_include],
        breaks = color_sequence[breaks_to_include],
        ylab = "Longitude",
        xlab = "Latitude",
        xlim = c(-176.5, -156.5),
        ylim = c(52, 62),
        main = 'Change in distribution',
        cex.main = 1.2,
        cex.lab = 1.1,
        cex.axis = 1.1)
  maps::map("worldHires",
            fill = T,
            col = "wheat4",
            add = T)
  image.plot(legend.only = T,
             col = my_color(n = color_levels)[col_to_include],
             legend.shrink = 0.2,
             smallplot = c(.73, .76, .25, .39),
             legend.cex = 0.6,
             axis.args = list(cex.axis = 0.8),
             legend.width = 0.5,
             legend.mar = 6,
             zlim = c(min(grid$diff, na.rm = T), max(grid$diff, na.rm = T)),
             legend.args = list("Predicted \n Change",
                                side = 2, cex = 0.8))
  plot(grid2$doy,
       grid2$pred,
       main = 'Change in phenology',
       type = 'l',
       ylim = range(c(grid2$pred_up,
                      grid2$pred2_up,
                      grid2$pred_lw,
                      grid2$pred2_lw)),
       xlim = c(60, 215),
       col = 'blue',
       lwd = 2,
       xlab = 'Day of the year',
       ylab = 'Egg density ln(n/10m2)',
       cex.lab = 1.1,
       cex.axis = 1.1,
       cex.main = 1.2)
  polygon(c(grid2$doy, rev(grid2$doy)),
          c(grid2$pred_lw, rev(grid2$pred_up)),
          col = alpha('blue', 0.2), 
          lty = 0)
  lines(grid2$doy,
        grid2$pred2,
        col = 'red',
        lwd = 2)
  polygon(c(grid2$doy, rev(grid2$doy)),
    c(grid2$pred2_lw, rev(grid2$pred2_up)),
    col = alpha('red', 0.2),
    lty = 0)
  legend(110,
         0,
         legend = c('Mean SST', '+ 1C'),
         col = c('blue', 'red'),
         lty = 1,
         bty = 'n',
         lwd = 2,
         cex = 0.8)
}

# Load fish data
pk_egg <- readRDS(here('data', 'pk_egg.rds'))
pk_larvae <- readRDS(here('data', 'pk_larvae.rds'))

# Load ROMS temperature means
roms_temps <- readRDS(here('data', 'roms_temps.rds'))
```

Base model:
Currently using only presence in a Gaussian GAM, may switch to Tweedie
With Gaussian, multiplication by binomial GAM required (two-stage)
```{r}
pk_egg_base <- gam(log(larvalcatchper10m2 + 1) ~ factor(year) +
                     s(doy) +
                     s(lon, lat),
                   data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_base)
```

Models with flexible phenology or geography
Note variable coefficient by year for each
```{r}
flex_phenology <- formula(log(larvalcatchper10m2 + 1) ~ factor(year) + 
                        s(lon, lat) + 
                        s(doy, by = factor(year)))
pk_egg_phenology <- gam(flex_phenology, data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_phenology)
AIC(pk_egg_phenology)
# AIC: 8708

flex_geography <- formula(log(larvalcatchper10m2 + 1) ~ factor(year) +
                        s(lon, lat, by = factor(year)) +
                        s(doy))
pk_egg_geography <- gam(flex_geography, data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_geography)
AIC(pk_egg_geography)
# AIC: 8390
```

Ratio for comparison between base and flexible models
Use to look at mean square error, base should have higher error
```{r}
ratio_pk_phenology <- (summary(pk_egg_base)$scale - 
                      summary(pk_egg_phenology)$scale) / summary(pk_egg_base)$scale
ratio_pk_phenology

ratio_pk_geography <- (summary(pk_egg_base)$scale - 
                      summary(pk_egg_geography)$scale) / summary(pk_egg_base)$scale
ratio_pk_geography
```

Add ROMS mean temperatures
Use a representative area with averaged values from Feb - April for each year
This can then be matched to the egg/larvae data for use in variable coefficient by SST 
instead of year as above

Phenology
Compare formulations with temp/salinity formulations
```{r}
# Add monthly temperature values from ROMS
pk_egg$mean_temp <- roms_temps$mean[match(pk_egg$year, roms_temps$year)]
range(pk_egg$mean_temp)

# This will again use the presence GAM from the two stage models
# create variable phenology model by adding doy*sst
pk_egg_month1 <- gam(log(larvalcatchper10m2 + 1) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       s(roms_temperature, k = 6) +
                       s(roms_salinity, k = 6),
                     data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_month1)
AIC(pk_egg_month1)
# Deviance explained: 45.2%
# R2 adj: 0.433
# AIC: 8688
par(mfrow = c(2, 2))
gam.check(pk_egg_month1)


pk_egg_month2 <- gam(log(larvalcatchper10m2 + 1) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       s(roms_temperature, k = 6) +
                       s(roms_salinity, k = 6) +
                       s(doy, by = mean_temp, k = 6),
                    data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_month2)
AIC(pk_egg_month2)
# Deviance explained: 45.9%
# R2 adj: 0.439
# AIC: 8673
par(mfrow = c(2, 2))
gam.check(pk_egg_month2)


pk_egg_month3 <- gam(log(larvalcatchper10m2 + 1) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       roms_temperature * roms_salinity,
                     data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_month3)
AIC(pk_egg_month3)
# Deviance explained: 40.9%
# R2 adj: 0.393
# AIC: 8826
par(mfrow = c(2, 2))
gam.check(pk_egg_month3)


pk_egg_month4 <- gam(log(larvalcatchper10m2 + 1) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       roms_temperature * roms_salinity +
                       s(roms_temperature, k = 6) +
                       s(roms_salinity, k = 6),
                     data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_month4)
AIC(pk_egg_month4)
# Deviance explained: 44.9%
# R2 adj: 0.432
# AIC: 8689
par(mfrow = c(2, 2))
gam.check(pk_egg_month4)


pk_egg_month5 <- gam(log(larvalcatchper10m2 + 1) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       roms_temperature * roms_salinity +
                       s(roms_temperature, k = 6) +
                       s(roms_salinity, k = 6) +
                       s(doy, by = mean_temp, k = 6),
                     data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_month5)
AIC(pk_egg_month5)
# Deviance explained: 45.2%
# R2 adj: 0.434
# AIC: 8683
par(mfrow = c(2, 2))
gam.check(pk_egg_month5)
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(pk_egg_month2, select = 1, main = "DOY")
plot(pk_egg_month2, select = 3, main = "Temperature")
plot(pk_egg_month2, select = 4, main = "Salinity")
plot(pk_egg_month2, select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_egg_roms_monthvc.jpg'), 
         height = 5, width = 10, units = 'in', res = 200)

par(mfrow = c(2, 2))
gam.check(pk_egg_month2)
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_egg_roms_monthvc_resid.jpg'), 
         height = 5, width = 10, units = 'in', res = 200)
```

Geography
Compare formulations with temp/salinity smooth terms
```{r}
# Create variable geography using space*sst

pk_egg_space2 <- gam(log(larvalcatchper10m2 + 1) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       s(roms_temperature, k = 6) +
                       s(roms_temperature, k = 6) +
                       s(lon, lat, by = mean_temp, k = 6),
                    data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_space2)
AIC(pk_egg_space2)
# Deviance explained: 45.4%
# R2 adj.: 0.432
# AIC: 8777
# Not very different from model with observational data


par(mfrow = c(2, 2))
plot(pk_egg_space, select = 1, main = "DOY")
plot(pk_egg_space, select = 2, main = "Location")
plot(pk_egg_space, select = 3, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_egg_roms_spacevc.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)

par(mfrow = c(2, 2))
gam.check(pk_egg_space)
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_egg_roms_spacevc_resid.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
```


VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_pk_month <- (summary(pk_egg_base)$scale - 
                         summary(pk_egg_month2)$scale) / summary(pk_egg_base)$scale
var_ratio_pk_month 

var_ratio_pk_space <- (summary(pk_egg_base)$scale - 
                         summary(pk_egg_space)$scale) / summary(pk_egg_base)$scale
var_ratio_pk_space
```

Plot the results of the GAMs that vary by year
This generates a three panel figure with a map, phenology curve, and MSE barplot
Grids:
```{r}
# Plot the results of the average geography and phenology and decrease of MSE
nlat = 80
nlon = 120
latd = seq(min(pk_egg$lat), max(pk_egg$lat), length.out = nlat)
lond = seq(min(pk_egg$lon), max(pk_egg$lon), length.out = nlon)
grid_pkegg_avg <- expand.grid(lond, latd)
names(grid_pkegg_avg) <- c('lon', 'lat')

# Geography grid
grid_pkegg_avg$dist <- NA
for (k in 1:nrow(grid_pkegg_avg)) {
  dist <-
    distance_function(grid_pkegg_avg$lat[k],
                      grid_pkegg_avg$lon[k],
                      pk_egg$lat,
                      pk_egg$lon)
  grid_pkegg_avg$dist[k] <- min(dist)
}
grid_pkegg_avg$year <- 2014
grid_pkegg_avg$doy <- median(pk_egg$doy)
grid_pkegg_avg$pred <- predict(pk_egg_base, newdata = grid_pkegg_avg)
grid_pkegg_avg$pred[grid_pkegg_avg$dist > 30000] <- NA

# Phenology grid
grid_pkegg_avg2 <-  data.frame('lon' = rep(-170, 100),
                                'lat' = rep(57, 100),
                                'doy' = seq(min(pk_egg$doy), 
                                            max(pk_egg$doy), 
                                            length = 100),
                                'year' = rep(2014, 100))
grid_pkegg_avg2$pred <- predict(pk_egg_base, newdata = grid_pkegg_avg2)
grid_pkegg_avg2$se <- predict(pk_egg_base, newdata = grid_pkegg_avg2, se = T)[[2]]
grid_pkegg_avg2$pred_up <- grid_pkegg_avg2$pred + 1.96 * grid_pkegg_avg2$se
grid_pkegg_avg2$pred_lw <- grid_pkegg_avg2$pred - 1.96 * grid_pkegg_avg2$se
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), mai = c(0.7, 0.6, 0.4, 0.4))
map_phenology(pk_egg, grid_pkegg_avg, grid_pkegg_avg2)

# Plot MSE as stacked bars, including year variability and SST variability
barplot(matrix(c(ratio_pk_geography,
                 var_ratio_pk_space,
                 ratio_pk_phenology,
                 var_ratio_pk_month) * 100, 
               ncol = 2,
               nrow = 2),
        ylab = 'Delta MSE (%)',
        names.arg = c('Variable distribution', 'Variable phenologies'),
        ylim = c(0, 35),
        main = 'Delta MSE',
        cex.lab = 1.1,
        cex.main = 1.2,
        cex.axis = 1.1,
        cex.names = 1.1,
        space = c(0, 1),
        beside = T,
        col = c('azure4', 'azure3'))
box()
legend("topright",
       legend = c('D-MSE|Year', 'D-MSE|SST'),
       bty = 'n',
       col = c('azure4', 'azure3'),
       pch = 15,
       pt.cex = 2.5,
       cex = 1.1)
dev.copy(jpeg,
         here('results/pollock_hindcast', 
              'pollock_egg_MSE.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
Spatial grid
```{r}
nlat = 80
nlon = 120
latd = seq(min(pk_egg$lat), max(pk_egg$lat), length.out = nlat)
lond = seq(min(pk_egg$lon), max(pk_egg$lon), length.out = nlon)
grid_vc_eggpk <- expand.grid(lond, latd)
names(grid_vc_eggpk) <- c('lon', 'lat')
grid_vc_eggpk$dist <- NA
for (k in 1:nrow(grid_vc_eggpk)) {
  dist <-  distance_function(grid_vc_eggpk$lat[k],
                             grid_vc_eggpk$lon[k],
                             pk_egg$lat,
                             pk_egg$lon)
  grid_vc_eggpk$dist[k] <- min(dist)
}
grid_vc_eggpk$year <- 2014
grid_vc_eggpk$doy <- median(pk_egg$doy)
grid_vc_eggpk$mean_temp <- mean(pk_egg$mean_temp)
grid_vc_eggpk$pred <- predict(pk_egg_geography, newdata = grid_vc_eggpk)
grid_vc_eggpk$se <- predict(pk_egg_geography, newdata = grid_vc_eggpk, se = T)[[2]]
grid_vc_eggpk$pred_up <- grid_vc_eggpk$pred + 1.96 * grid_vc_eggpk$se
grid_vc_eggpk$pred_lw <- grid_vc_eggpk$pred - 1.96 * grid_vc_eggpk$se
grid_vc_eggpk$pred[grid_vc_eggpk$dist > 30000] <- NA
grid_vc_eggpk$mean_temp <- mean(pk_egg$mean_temp) + 1
grid_vc_eggpk$pred2 <- predict(pk_egg_space, newdata = grid_vc_eggpk)
grid_vc_eggpk$se2 <- predict(pk_egg_space, newdata = grid_vc_eggpk, se = T)[[2]]
grid_vc_eggpk$pred2_up <- grid_vc_eggpk$pred2 + 1.96 * grid_vc_eggpk$se2
grid_vc_eggpk$pred2_lw <- grid_vc_eggpk$pred2 - 1.96 * grid_vc_eggpk$se2
grid_vc_eggpk$diff <- grid_vc_eggpk$pred2 - grid_vc_eggpk$pred
grid_vc_eggpk$sig_pos <- c(grid_vc_eggpk$pred2_lw > grid_vc_eggpk$pred_up)
grid_vc_eggpk$sig_neg <- c(grid_vc_eggpk$pred2_up < grid_vc_eggpk$pred_lw)
grid_vc_eggpk$pos_diff <- grid_vc_eggpk$diff * grid_vc_eggpk$sig_pos
grid_vc_eggpk$neg_diff <- grid_vc_eggpk$diff * grid_vc_eggpk$sig_neg
max_slope <- max(grid_vc_eggpk$diff, na.rm = T)
```

Temporal grid
```{r}
grid_vc_eggpk2 <-  data.frame('lon' = rep(-170, 100),
                           'lat' = rep(57, 100),
                           'doy' = seq(min(pk_egg$doy), 
                                       max(pk_egg$doy), 
                                       length = 100),
                           mean_temp = rep(mean(pk_egg$mean_temp), 100),
                           'year' = rep(2014, 100))
grid_vc_eggpk2$pred <- predict(pk_egg_month2, newdata = grid_vc_eggpk2)
grid_vc_eggpk2$se <- predict(pk_egg_month2, newdata = grid_vc_eggpk2, se = T)[[2]]
grid_vc_eggpk2$pred_up <- grid_vc_eggpk2$pred + 1.96 * grid_vc_eggpk2$se
grid_vc_eggpk2$pred_lw <- grid_vc_eggpk2$pred - 1.96 * grid_vc_eggpk2$se
grid_vc_eggpk2$mean_temp <- mean(pk_egg$mean_temp) + 1
grid_vc_eggpk2$pred2 <- predict(pk_egg_month2, newdata = grid_vc_eggpk2)
grid_vc_eggpk2$se2 <- predict(pk_egg_month2, newdata = grid_vc_eggpk2, se = T)[[2]]
grid_vc_eggpk2$pred2_up <- grid_vc_eggpk2$pred2 + 1.96 * grid_vc_eggpk2$se2
grid_vc_eggpk2$pred2_lw <- grid_vc_eggpk2$pred2 - 1.96 * grid_vc_eggpk2$se2
```

Create two panel plot
Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 2), 
    mai = c(0.8, 0.9, 0.5, 0.5))
map_vc(pk_egg, grid_vc_eggpk, grid_vc_eggpk2)
dev.copy(jpeg,
         here('results/pollock_hindcast',
         'pollock_egg_vc.jpg'),
         height = 3.5,
         width = 8,
         res = 200,
         units = 'in')
```
