---
title: "Hindcast ROMS variables"
author: "Rebecca Howard"
date: "9/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(marmap)
library(maps)
library(mapdata)
library(fields)
library(here)
library(mgcv)
library(tidyr)
library(dplyr)
library(ggplot2)
library(raster)
library(rgdal)
library(viridis)
library(RColorBrewer)
source(here('code/functions', 'vis_gam_COLORS.R'))
source(here('code/functions', 'distance_function.R'))
source(here('code/functions', 'gam_selection.R'))
source(here('code/functions', 'map_vc.R'))
source(here('code/functions', 'map_phenology.R'))
source(here('code/functions', 'threepanel_base_mse.R'))
source(here('code/functions', 'vc_grids.R'))
source(here('code/functions', 'plot_variable.R'))
source(here('code/functions', 'location_plot.R'))
```

```{r}
# Load ROMS temperature means
roms_temps <- readRDS(here('data', 'roms_temps.rds'))

# Load fish data
pk_egg <- as.data.frame(filter((readRDS(here('data', 'pk_egg.rds'))),
                 lat >= 52 & lat <= 62,
                  lon >= -176.5 & lon <= -156.5))
pk_egg$mean_temp <- roms_temps$mean[match(pk_egg$year, roms_temps$year)]
pk_egg$catch <- pk_egg$catch

pk_larvae <- as.data.frame(filter(readRDS(here('data', 'pk_larvae.rds')),
                 lat >= 52 & lat <= 62,
                  lon >= -176.5 & lon <= -156.5))
pk_larvae$mean_temp <- roms_temps$mean[match(pk_larvae$year, roms_temps$year)]
pk_larvae$catch <- pk_larvae$catch

fhs_egg <- as.data.frame(filter(readRDS(here('data', 'fhs_egg.rds')),
                 lat >= 52 & lat <= 62,
                  lon >= -176.5 & lon <= -156.5))
fhs_egg$mean_temp <- roms_temps$mean[match(fhs_egg$year, roms_temps$year)]
fhs_egg$catch <- fhs_egg$catch

fhs_larvae <- as.data.frame(filter(readRDS(here('data', 'fhs_larvae.rds')),
                 lat >= 52 & lat <= 62,
                  lon >= -176.5 & lon <= -156.5))
fhs_larvae$mean_temp <- roms_temps$mean[match(fhs_larvae$year, roms_temps$year)]
fhs_larvae$catch <- fhs_larvae$catch

yfs_egg <- as.data.frame(filter(readRDS(here('data', 'yfs_egg.rds')),
                 lat >= 52 & lat <= 62,
                  lon >= -176.5 & lon <= -156.5))
yfs_egg$mean_temp <- roms_temps$mean[match(yfs_egg$year, roms_temps$year)]
yfs_egg$catch <- yfs_egg$catch

yfs_larvae <- as.data.frame(filter(readRDS(here('data', 'yfs_larvae.rds')),
                 lat >= 52 & lat <= 62,
                  lon >= -176.5 & lon <= -156.5))
yfs_larvae$mean_temp <- roms_temps$mean[match(yfs_larvae$year, roms_temps$year)]
yfs_larvae$catch <- yfs_larvae$catch

akp_egg <- as.data.frame(filter(readRDS(here('data', 'akp_egg.rds')),
                 lat >= 52 & lat <= 62,
                  lon >= -176.5 & lon <= -156.5))
akp_egg$mean_temp <- roms_temps$mean[match(akp_egg$year, roms_temps$year)]
akp_egg$catch <- akp_egg$catch

akp_larvae <- as.data.frame(filter(readRDS(here('data', 'akp_larvae.rds')),
                 lat >= 52 & lat <= 62,
                  lon >= -176.5 & lon <= -156.5))
akp_larvae$mean_temp <- roms_temps$mean[match(akp_larvae$year, roms_temps$year)]
akp_larvae$catch <- akp_larvae$catch

# Colors
contour_col <- rgb(0, 0, 255, max = 255, alpha = 0, names = "white")

jet.colors = colorRampPalette(rev(c("#FFFFCC", "#FBF2A8", "#F9E585",
                                    "#F5D363", "#EFBA55", "#EAA352",
                                    "#E68C51", "#E0754F", "#D75C4D",
                                    "#BB4A48", "#994240", "#763931", 
                                    "#542D20", "#352311", "#191900")))
```

Base models: Currently using only presence in a Gaussian GAM, may switch to Tweedie With Gaussian, multiplication by binomial GAM required (two-stage)
```{r}
# Gaussian
pk_egg_base <- gam(log(catch) ~ factor(year) +
                     s(doy) +
                     s(lon, lat),
                   data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_base)

pk_larvae_base <- gam(log(catch) ~ factor(year) +
                     s(doy) +
                     s(lon, lat),
                   data = pk_larvae[pk_larvae$larvalcatchper10m2 > 0, ])
summary(pk_larvae_base)
```
#### Pollock
### Gaussian
## Eggs

Models with flexible phenology or geography 
Note variable coefficient by year for each
```{r}
pk_egg_phenology <- gam(log(catch) ~ factor(year) + 
                          s(doy) +
                          s(lon, lat) + 
                          s(doy, by = factor(year)), 
                        data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_phenology)
AIC(pk_egg_phenology)
# AIC: 8713

pk_egg_geography <- gam(log(catch) ~ factor(year) +
                          s(lon, lat) +
                          s(lon, lat, by = factor(year)) +
                          s(doy), 
                        data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_geography)
AIC(pk_egg_geography)
# AIC: 8369
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_pkegg_phenology <- (summary(pk_egg_base)$scale - 
                      summary(pk_egg_phenology)$scale) / summary(pk_egg_base)$scale
ratio_pkegg_phenology

ratio_pkegg_geography <- (summary(pk_egg_base)$scale - 
                      summary(pk_egg_geography)$scale) / summary(pk_egg_base)$scale
ratio_pkegg_geography
```

Add ROMS mean temperatures Use a representative area with averaged values from Feb - April for each year This can then be matched to the egg/larvae data for use in variable coefficient by SST instead of year as above

Phenology Compare formulations with temp/salinity formulations
```{r}
# Add monthly temperature values from ROMS
pk_egg$mean_temp <- roms_temps$mean[match(pk_egg$year, roms_temps$year)]
range(pk_egg$mean_temp)

# This will again use the presence GAM from the two stage models
# create variable phenology model by adding doy*sst
pk_egg_month1 <- gam(log(catch) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       s(roms_temperature, k = 6) +
                       s(roms_salinity, k = 6),
                     data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_month1)
AIC(pk_egg_month1)
# Deviance explained: 44.8%
# R2 adj: 0.431
# AIC: 8691
par(mfrow = c(2, 2))
gam.check(pk_egg_month1)


pk_egg_month2 <- gam(log(catch) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       s(roms_temperature, k = 6) +
                       s(roms_salinity, k = 6) +
                       s(doy, by = mean_temp, k = 6),
                    data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_month2)
AIC(pk_egg_month2)
# Deviance explained: 45.1%
# R2 adj: 0.433
# AIC: 8685
par(mfrow = c(2, 2))
gam.check(pk_egg_month2)


pk_egg_month3 <- gam(log(catch) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       s(roms_temperature, roms_salinity),
                     data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_month3)
AIC(pk_egg_month3)
# Deviance explained: 40.9%
# R2 adj: 0.435
# AIC: 8688
par(mfrow = c(2, 2))
gam.check(pk_egg_month3)


pk_egg_month4 <- gam(log(catch) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       s(roms_temperature, roms_salinity) +
                       s(roms_temperature, k = 6) +
                       s(roms_salinity, k = 6),
                     data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_month4)
AIC(pk_egg_month4)
# Deviance explained: 45.9%
# R2 adj: 0.438
# AIC: 8676
par(mfrow = c(2, 2))
gam.check(pk_egg_month4)


pk_egg_month5 <- gam(log(catch) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       s(roms_temperature, roms_salinity) +
                       s(roms_temperature, k = 6) +
                       s(roms_salinity, k = 6) +
                       s(doy, by = mean_temp, k = 6),
                     data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_month5)
AIC(pk_egg_month5)
# Deviance explained: 46.2%
# R2 adj: 0.441
# AIC: 8670
par(mfrow = c(2, 2))
gam.check(pk_egg_month5)
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(pk_egg_month5, select = 1, main = "DOY")
plot(pk_egg_month5, select = 3, main = "Temperature")
plot(pk_egg_month5, select = 4, main = "Salinity")
plot(pk_egg_month5, select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_egg_roms_monthvc.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(pk_egg_month5)
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_egg_roms_monthvc_resid.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Geography Compare formulations with temp/salinity smooth terms
```{r}
# Create variable geography using space*sst
# Compare to the first three phenology models for model selection
pk_egg_space1 <- gam(log(catch) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       s(roms_temperature, k = 6) +
                       s(roms_salinity, k = 6) +
                       s(lon, lat, by = mean_temp, k = 6),
                    data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_space1)
AIC(pk_egg_space1)
# Deviance explained: 45.3%
# R2 adj.: 0.435
# AIC: 8679
par(mfrow = c(2, 2))
gam.check(pk_egg_space1)


pk_egg_space2 <- gam(log(catch) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       s(roms_temperature, roms_salinity) +
                       s(roms_temperature, k = 6) +
                       s(roms_salinity, k = 6) +
                       s(lon, lat, by = mean_temp, k = 6),
                    data = pk_egg[pk_egg$larvalcatchper10m2 > 0, ])
summary(pk_egg_space2)
AIC(pk_egg_space2)
# Deviance explained: 46.3%
# R2 adj.: 0.442
# AIC: 8666
par(mfrow = c(2, 2))
gam.check(pk_egg_space2)
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(pk_egg_space2, select = 1, main = "DOY")
plot(pk_egg_space2, select = 3, main = "Temperature")
plot(pk_egg_space2, select = 4, main = "Salinity")
plot(pk_egg_space2, select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_egg_roms_spacevc.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(pk_egg_space2)
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_egg_roms_spacevc_resid.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_pkegg_month <- (summary(pk_egg_base)$scale - 
                         summary(pk_egg_month2)$scale) / summary(pk_egg_base)$scale
var_ratio_pkegg_month 

var_ratio_pkegg_space <- (summary(pk_egg_base)$scale - 
                         summary(pk_egg_space2)$scale) / summary(pk_egg_base)$scale
var_ratio_pkegg_space
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
# Plot the results of the average geography and phenology and decrease of MSE
nlat = 80
nlon = 120
latd = seq(min(pk_egg$lat), max(pk_egg$lat), length.out = nlat)
lond = seq(min(pk_egg$lon), max(pk_egg$lon), length.out = nlon)
grid_pkegg_avg <- expand.grid(lond, latd)
names(grid_pkegg_avg) <- c('lon', 'lat')

# Geography grid
grid_pkegg_avg$dist <- NA
for (k in 1:nrow(grid_pkegg_avg)) {
  dist <-
    distance_function(grid_pkegg_avg$lat[k],
                      grid_pkegg_avg$lon[k],
                      pk_egg$lat,
                      pk_egg$lon)
  grid_pkegg_avg$dist[k] <- min(dist)
}
grid_pkegg_avg$year <- 2014
grid_pkegg_avg$doy <- median(pk_egg$doy)
grid_pkegg_avg$pred <- predict(pk_egg_base, newdata = grid_pkegg_avg)
grid_pkegg_avg$pred[grid_pkegg_avg$dist > 30000] <- NA

# Phenology grid
grid_pkegg_avg2 <-  data.frame('lon' = rep(-170, 100),
                                'lat' = rep(57, 100),
                                'doy' = seq(min(pk_egg$doy), 
                                            max(pk_egg$doy), 
                                            length = 100),
                                'year' = rep(2014, 100))
grid_pkegg_avg2$pred <- predict(pk_egg_base, newdata = grid_pkegg_avg2)
grid_pkegg_avg2$se <- predict(pk_egg_base, newdata = grid_pkegg_avg2, se = T)[[2]]
grid_pkegg_avg2$pred_up <- grid_pkegg_avg2$pred + 1.96 * grid_pkegg_avg2$se
grid_pkegg_avg2$pred_lw <- grid_pkegg_avg2$pred - 1.96 * grid_pkegg_avg2$se
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(pk_egg, grid_pkegg_avg, grid_pkegg_avg2)

# Plot MSE as stacked bars, including year variability and SST variability
barplot(matrix(c(ratio_pkegg_geography,
                 var_ratio_pkegg_space,
                 ratio_pkegg_phenology,
                 var_ratio_pkegg_month) * 100, 
               ncol = 2,
               nrow = 2),
        ylab = 'Delta MSE (%)',
        names.arg = c('Variable distribution', 'Variable phenologies'),
        ylim = c(0, 35),
        main = 'Delta MSE',
        cex.lab = 1.1,
        cex.main = 1.2,
        cex.axis = 1.1,
        cex.names = 1.1,
        space = c(0, 1),
        beside = T,
        col = c('azure4', 'azure3'))
box()
legend("topright",
       legend = c('D-MSE|Year', 'D-MSE|SST'),
       bty = 'n',
       col = c('azure4', 'azure3'),
       pch = 15,
       pt.cex = 2.5,
       cex = 1.1)
dev.copy(jpeg,
         here('results/pollock_hindcast', 
              'pollock_egg_MSE.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs Spatial grid
```{r}
nlat = 80
nlon = 120
latd = seq(min(pk_egg$lat), max(pk_egg$lat), length.out = nlat)
lond = seq(min(pk_egg$lon), max(pk_egg$lon), length.out = nlon)
grid_vc_eggpk <- expand.grid(lond, latd)
names(grid_vc_eggpk) <- c('lon', 'lat')
grid_vc_eggpk$dist <- NA
for (k in 1:nrow(grid_vc_eggpk)) {
  dist <-  distance_function(grid_vc_eggpk$lat[k],
                             grid_vc_eggpk$lon[k],
                             pk_egg$lat,
                             pk_egg$lon)
  grid_vc_eggpk$dist[k] <- min(dist)
}
grid_vc_eggpk$year <- 2014
grid_vc_eggpk$doy <- median(pk_egg$doy)
grid_vc_eggpk$mean_temp <- mean(pk_egg$mean_temp)
grid_vc_eggpk$roms_temperature <- mean(pk_egg$roms_temperature, na.rm = T)
grid_vc_eggpk$roms_salinity <- mean(pk_egg$roms_salinity, na.rm = T)
grid_vc_eggpk$pred <- predict(pk_egg_space2, newdata = grid_vc_eggpk)
grid_vc_eggpk$se <- predict(pk_egg_space2, newdata = grid_vc_eggpk, se = T)[[2]]
grid_vc_eggpk$pred_up <- grid_vc_eggpk$pred + 1.96 * grid_vc_eggpk$se
grid_vc_eggpk$pred_lw <- grid_vc_eggpk$pred - 1.96 * grid_vc_eggpk$se
grid_vc_eggpk$pred[grid_vc_eggpk$dist > 30000] <- NA
grid_vc_eggpk$mean_temp <- mean(pk_egg$mean_temp) + 1
grid_vc_eggpk$pred2 <- predict(pk_egg_space2, newdata = grid_vc_eggpk)
grid_vc_eggpk$se2 <- predict(pk_egg_space2, newdata = grid_vc_eggpk, se = T)[[2]]
grid_vc_eggpk$pred2_up <- grid_vc_eggpk$pred2 + 1.96 * grid_vc_eggpk$se2
grid_vc_eggpk$pred2_lw <- grid_vc_eggpk$pred2 - 1.96 * grid_vc_eggpk$se2
grid_vc_eggpk$diff <- grid_vc_eggpk$pred2 - grid_vc_eggpk$pred
grid_vc_eggpk$sig_pos <- c(grid_vc_eggpk$pred2_lw > grid_vc_eggpk$pred_up)
grid_vc_eggpk$sig_neg <- c(grid_vc_eggpk$pred2_up < grid_vc_eggpk$pred_lw)
grid_vc_eggpk$pos_diff <- grid_vc_eggpk$diff * grid_vc_eggpk$sig_pos
grid_vc_eggpk$neg_diff <- grid_vc_eggpk$diff * grid_vc_eggpk$sig_neg
max_slope <- max(grid_vc_eggpk$diff, na.rm = T)
```

Temporal grid
```{r}
grid_vc_eggpk2 <-  data.frame('lon' = rep(-170, 100),
                           'lat' = rep(57, 100),
                           'doy' = seq(min(pk_egg$doy), 
                                       max(pk_egg$doy), 
                                       length = 100),
                           mean_temp = rep(mean(pk_egg$mean_temp), 100),
                           'year' = rep(2014, 100),
                           'roms_temperature' = rep(mean(pk_egg$roms_temperature, na.rm = T), 100),
                           'roms_salinity' = rep(mean(pk_egg$roms_salinity, na.rm = T), 100))
grid_vc_eggpk2$pred <- predict(pk_egg_month5, newdata = grid_vc_eggpk2)
grid_vc_eggpk2$se <- predict(pk_egg_month5, newdata = grid_vc_eggpk2, se = T)[[2]]
grid_vc_eggpk2$pred_up <- grid_vc_eggpk2$pred + 1.96 * grid_vc_eggpk2$se
grid_vc_eggpk2$pred_lw <- grid_vc_eggpk2$pred - 1.96 * grid_vc_eggpk2$se
grid_vc_eggpk2$mean_temp <- mean(pk_egg$mean_temp) + 1
grid_vc_eggpk2$pred2 <- predict(pk_egg_month5, newdata = grid_vc_eggpk2)
grid_vc_eggpk2$se2 <- predict(pk_egg_month5, newdata = grid_vc_eggpk2, se = T)[[2]]
grid_vc_eggpk2$pred2_up <- grid_vc_eggpk2$pred2 + 1.96 * grid_vc_eggpk2$se2
grid_vc_eggpk2$pred2_lw <- grid_vc_eggpk2$pred2 - 1.96 * grid_vc_eggpk2$se2
```

Create two panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc(pk_egg, pk_egg_space2, grid_vc_eggpk, grid_vc_eggpk2, "Spatial GAM Hindcast")
  legend(130,
         11,
         legend = c('Mean SST', '+ 1C'),
         col = c('#000000', '#E69F00'),
         lty = 1,
         bty = 'n',
         lwd = 2,
         cex = 0.6)
dev.copy(jpeg,
         here('results/pollock_hindcast',
         'pollock_egg_best.jpg'),
         height = 3.5,
         width = 8,
         res = 200,
         units = 'in')
dev.off()
```

## Larvae
Base model:
Currently using only presence in a Gaussian GAM, may switch to Tweedie
With Gaussian, multiplication by binomial GAM required (two-stage)
```{r}
# Gaussian
pk_larvae_base <- gam(log(catch) ~ factor(year) +
                     s(doy) +
                     s(lon, lat),
                   data = pk_larvae[pk_larvae$larvalcatchper10m2 > 0, ])
summary(pk_larvae_base)

# Tweedie
pk_larvae_base_t <- gam(catch ~ factor(year) +
                       s(lon, lat) + 
                       s(doy),
                    data = pk_larvae,
                    family = tw(link = 'log'),
                    method = 'REML')
summary(pk_larvae_base_t)
```

Models with flexible phenology or geography
Note variable coefficient by year for each
```{r}
pk_larvae_phenology <- gam(log(catch) ~ factor(year) + 
                             s(lon, lat) +
                             s(doy) +
                             s(doy, by = factor(year)), 
                           data = pk_larvae[pk_larvae$larvalcatchper10m2 > 0, ])
summary(pk_larvae_phenology)
AIC(pk_larvae_phenology)
# AIC: 7501

pk_larvae_geography <- gam(log(catch) ~ factor(year) +
                             s(lon, lat) +
                             s(lon, lat, by = factor(year)) +
                             s(doy), 
                           data = pk_larvae[pk_larvae$larvalcatchper10m2 > 0, ])
summary(pk_larvae_geography)
AIC(pk_larvae_geography)
# AIC: 6694
```

Ratio for comparison between base and flexible models
Use to look at mean square error, base should have higher error
```{r}
ratio_pklarvae_phenology <- (summary(pk_larvae_base)$scale - 
                      summary(pk_larvae_phenology)$scale) / summary(pk_larvae_base)$scale
ratio_pklarvae_phenology

ratio_pklarvae_geography <- (summary(pk_larvae_base)$scale - 
                      summary(pk_larvae_geography)$scale) / summary(pk_larvae_base)$scale
ratio_pklarvae_geography 
```

Add ROMS mean temperatures
Use a representative area with averaged values from Feb - April for each year
This can then be matched to the egg/larvae data for use in variable coefficient by SST 
instead of year as above

Phenology
Compare formulations with temp/salinity formulations
```{r}
# Add monthly temperature values from ROMS
pk_larvae$mean_temp <- roms_temps$mean[match(pk_larvae$year, roms_temps$year)]
range(pk_larvae$mean_temp)

# This will again use the presence GAM from the two stage models
# create variable phenology model by adding doy*sst
pk_larvae_month1 <- gam(log(catch) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       s(roms_temperature, k = 6) +
                       s(roms_salinity, k = 6),
                     data = pk_larvae[pk_larvae$larvalcatchper10m2 > 0, ])
summary(pk_larvae_month1)
AIC(pk_larvae_month1)
# Deviance explained: 48.1%
# R2 adj: 0.464
# AIC: 7523
par(mfrow = c(2, 2))
gam.check(pk_larvae_month1)


pk_larvae_month2 <- gam(log(catch) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       s(roms_temperature, k = 6) +
                       s(roms_salinity, k = 6) +
                       s(doy, by = mean_temp, k = 6),
                    data = pk_larvae[pk_larvae$larvalcatchper10m2 > 0, ])
summary(pk_larvae_month2)
AIC(pk_larvae_month2)
# Deviance explained: 48.6%
# R2 adj: 0.468
# AIC: 7513
par(mfrow = c(2, 2))
gam.check(pk_larvae_month2)


pk_larvae_month3 <- gam(log(catch) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       s(roms_temperature, roms_salinity),
                     data = pk_larvae[pk_larvae$larvalcatchper10m2 > 0, ])
summary(pk_larvae_month3)
AIC(pk_larvae_month3)
# Deviance explained: 45%
# R2 adj: 0.433
# AIC: 7631
par(mfrow = c(2, 2))
gam.check(pk_larvae_month3)


pk_larvae_month4 <- gam(log(catch) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       s(roms_temperature, roms_salinity) +
                       s(roms_temperature, k = 6) +
                       s(roms_salinity, k = 6),
                     data = pk_larvae[pk_larvae$larvalcatchper10m2 > 0, ])
summary(pk_larvae_month4)
AIC(pk_larvae_month4)
# Deviance explained: 48.3%
# R2 adj: 0.465
# AIC: 7518
par(mfrow = c(2, 2))
gam.check(pk_larvae_month4)


pk_larvae_month5 <- gam(log(catch) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       s(roms_temperature, roms_salinity) +
                       s(roms_temperature, k = 6) +
                       s(roms_salinity, k = 6) +
                       s(doy, by = mean_temp, k = 6),
                     data = pk_larvae[pk_larvae$larvalcatchper10m2 > 0, ])
summary(pk_larvae_month5)
AIC(pk_larvae_month5)
# Deviance explained: 48.7%
# R2 adj: 0.469
# AIC: 7510
par(mfrow = c(2, 2))
gam.check(pk_larvae_month5)
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(pk_larvae_month5, select = 1, main = "DOY")
plot(pk_larvae_month5, select = 3, main = "Temperature")
plot(pk_larvae_month5, select = 4, main = "Salinity")
plot(pk_larvae_month5, select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_larvae_roms_monthvc.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(pk_larvae_month5)
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_larvae_roms_monthvc_resid.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Geography
Compare formulations with temp/salinity smooth terms
```{r}
# Create variable geography using space*sst
# Compare to the first three phenology models for model selection
pk_larvae_space1 <- gam(log(catch) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       s(roms_temperature, k = 6) +
                       s(roms_salinity, k = 6) +
                       s(lon, lat, by = mean_temp, k = 6),
                    data = pk_larvae[pk_larvae$larvalcatchper10m2 > 0, ])
summary(pk_larvae_space1)
AIC(pk_larvae_space1)
# Deviance explained: 50.7%
# R2 adj.: 0.489
# AIC: 7431
par(mfrow = c(2, 2))
gam.check(pk_larvae_space1)


pk_larvae_space2 <- gam(log(catch) ~ factor(year) +
                       s(doy, k = 8) +
                       s(lon, lat) +
                       s(roms_temperature, roms_salinity) +
                       s(roms_temperature, k = 6) +
                       s(roms_salinity, k = 6) +
                       s(lon, lat, by = mean_temp, k = 6),
                    data = pk_larvae[pk_larvae$larvalcatchper10m2 > 0, ])
summary(pk_larvae_space2)
AIC(pk_larvae_space2)
# Deviance explained: 50.7%
# R2 adj.: 0.489
# AIC: 7431
par(mfrow = c(2, 2))
gam.check(pk_larvae_space2)
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(pk_larvae_space2, select = 1, main = "DOY")
plot(pk_larvae_space2, select = 3, main = "Temperature")
plot(pk_larvae_space2, select = 4, main = "Salinity")
plot(pk_larvae_space2, select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_larvae_roms_spacevc.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(pk_larvae_space2)
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_larvae_roms_spacevc_resid.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```


VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_pklarvae_month <- (summary(pk_larvae_base)$scale - 
                         summary(pk_larvae_month2)$scale) / summary(pk_larvae_base)$scale
var_ratio_pklarvae_month 

var_ratio_pklarvae_space <- (summary(pk_larvae_base)$scale - 
                         summary(pk_larvae_space2)$scale) / summary(pk_larvae_base)$scale
var_ratio_pklarvae_space
```

Plot the results of the GAMs that vary by year
This generates a three panel figure with a map, phenology curve, and MSE barplot
Grids:
```{r}
# Plot the results of the average geography and phenology and decrease of MSE
nlat = 80
nlon = 120
latd = seq(min(pk_larvae$lat), max(pk_larvae$lat), length.out = nlat)
lond = seq(min(pk_larvae$lon), max(pk_larvae$lon), length.out = nlon)
grid_pklarvae_avg <- expand.grid(lond, latd)
names(grid_pklarvae_avg) <- c('lon', 'lat')

# Geography grid
grid_pklarvae_avg$dist <- NA
for (k in 1:nrow(grid_pklarvae_avg)) {
  dist <-
    distance_function(grid_pklarvae_avg$lat[k],
                      grid_pklarvae_avg$lon[k],
                      pk_larvae$lat,
                      pk_larvae$lon)
  grid_pklarvae_avg$dist[k] <- min(dist)
}
grid_pklarvae_avg$year <- 2014
grid_pklarvae_avg$doy <- median(pk_larvae$doy)
grid_pklarvae_avg$pred <- predict(pk_larvae_base, newdata = grid_pklarvae_avg)
grid_pklarvae_avg$pred[grid_pklarvae_avg$dist > 30000] <- NA

# Phenology grid
grid_pklarvae_avg2 <-  data.frame('lon' = rep(-170, 100),
                                'lat' = rep(57, 100),
                                'doy' = seq(min(pk_larvae$doy), 
                                            max(pk_larvae$doy), 
                                            length = 100),
                                'year' = rep(2014, 100))
grid_pklarvae_avg2$pred <- predict(pk_larvae_base, newdata = grid_pklarvae_avg2)
grid_pklarvae_avg2$se <- predict(pk_larvae_base, newdata = grid_pklarvae_avg2, se = T)[[2]]
grid_pklarvae_avg2$pred_up <- grid_pklarvae_avg2$pred + 1.96 * grid_pklarvae_avg2$se
grid_pklarvae_avg2$pred_lw <- grid_pklarvae_avg2$pred - 1.96 * grid_pklarvae_avg2$se
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), mai = c(0.7, 0.6, 0.4, 0.4))
map_phenology(pk_larvae, grid_pklarvae_avg, grid_pklarvae_avg2)

# Plot MSE as stacked bars, including year variability and SST variability
barplot(matrix(c(ratio_pklarvae_geography,
                 var_ratio_pklarvae_space,
                 ratio_pklarvae_phenology,
                 var_ratio_pklarvae_month) * 100, 
               ncol = 2,
               nrow = 2),
        ylab = 'Delta MSE (%)',
        names.arg = c('Variable distribution', 'Variable phenologies'),
        ylim = c(0, 50),
        main = 'Delta MSE',
        cex.lab = 1.1,
        cex.main = 1.2,
        cex.axis = 1.1,
        cex.names = 1.1,
        space = c(0, 1),
        beside = T,
        col = c('azure4', 'azure3'))
box()
legend("topright",
       legend = c('D-MSE|Year', 'D-MSE|SST'),
       bty = 'n',
       col = c('azure4', 'azure3'),
       pch = 15,
       pt.cex = 2.5,
       cex = 1.1)
dev.copy(jpeg,
         here('results/pollock_hindcast', 
              'pollock_larvae_MSE.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
Spatial grid
```{r}
nlat = 80
nlon = 120
latd = seq(min(pk_larvae$lat), max(pk_larvae$lat), length.out = nlat)
lond = seq(min(pk_larvae$lon), max(pk_larvae$lon), length.out = nlon)
grid_vc_larvaepk <- expand.grid(lond, latd)
names(grid_vc_larvaepk) <- c('lon', 'lat')
grid_vc_larvaepk$dist <- NA
for (k in 1:nrow(grid_vc_larvaepk)) {
  dist <-  distance_function(grid_vc_larvaepk$lat[k],
                             grid_vc_larvaepk$lon[k],
                             pk_larvae$lat,
                             pk_larvae$lon)
  grid_vc_larvaepk$dist[k] <- min(dist)
}
grid_vc_larvaepk$year <- 2014
grid_vc_larvaepk$doy <- median(pk_larvae$doy)
grid_vc_larvaepk$mean_temp <- mean(pk_larvae$mean_temp)
grid_vc_larvaepk$roms_temperature <- mean(pk_larvae$roms_temperature, na.rm = T)
grid_vc_larvaepk$roms_salinity <- mean(pk_larvae$roms_salinity, na.rm = T)
grid_vc_larvaepk$pred <- predict(pk_larvae_space2, newdata = grid_vc_larvaepk)
grid_vc_larvaepk$se <- predict(pk_larvae_space2, newdata = grid_vc_larvaepk, se = T)[[2]]
grid_vc_larvaepk$pred_up <- grid_vc_larvaepk$pred + 1.96 * grid_vc_larvaepk$se
grid_vc_larvaepk$pred_lw <- grid_vc_larvaepk$pred - 1.96 * grid_vc_larvaepk$se
grid_vc_larvaepk$pred[grid_vc_larvaepk$dist > 30000] <- NA
grid_vc_larvaepk$mean_temp <- mean(pk_larvae$mean_temp) + 1
grid_vc_larvaepk$pred2 <- predict(pk_larvae_space2, newdata = grid_vc_larvaepk)
grid_vc_larvaepk$se2 <- predict(pk_larvae_space2, newdata = grid_vc_larvaepk, se = T)[[2]]
grid_vc_larvaepk$pred2_up <- grid_vc_larvaepk$pred2 + 1.96 * grid_vc_larvaepk$se2
grid_vc_larvaepk$pred2_lw <- grid_vc_larvaepk$pred2 - 1.96 * grid_vc_larvaepk$se2
grid_vc_larvaepk$diff <- grid_vc_larvaepk$pred2 - grid_vc_larvaepk$pred
grid_vc_larvaepk$sig_pos <- c(grid_vc_larvaepk$pred2_lw > grid_vc_larvaepk$pred_up)
grid_vc_larvaepk$sig_neg <- c(grid_vc_larvaepk$pred2_up < grid_vc_larvaepk$pred_lw)
grid_vc_larvaepk$pos_diff <- grid_vc_larvaepk$diff * grid_vc_larvaepk$sig_pos
grid_vc_larvaepk$neg_diff <- grid_vc_larvaepk$diff * grid_vc_larvaepk$sig_neg
max_slope <- max(grid_vc_larvaepk$diff, na.rm = T)
```

Temporal grid
```{r}
grid_vc_larvaepk2 <-  data.frame('lon' = rep(-170, 100),
                           'lat' = rep(57, 100),
                           'doy' = seq(min(pk_larvae$doy), 
                                       max(pk_larvae$doy), 
                                       length = 100),
                           mean_temp = rep(mean(pk_larvae$mean_temp), 100),
                           'year' = rep(2014, 100),
                           'roms_temperature' = rep(mean(pk_larvae$roms_temperature, na.rm = T), 100),
                           'roms_salinity' = rep(mean(pk_larvae$roms_salinity, na.rm = T), 100))
grid_vc_larvaepk2$pred <- predict(pk_larvae_month5, newdata = grid_vc_larvaepk2)
grid_vc_larvaepk2$se <- predict(pk_larvae_month5, newdata = grid_vc_larvaepk2, se = T)[[2]]
grid_vc_larvaepk2$pred_up <- grid_vc_larvaepk2$pred + 1.96 * grid_vc_larvaepk2$se
grid_vc_larvaepk2$pred_lw <- grid_vc_larvaepk2$pred - 1.96 * grid_vc_larvaepk2$se
grid_vc_larvaepk2$mean_temp <- mean(pk_larvae$mean_temp) + 1
grid_vc_larvaepk2$pred2 <- predict(pk_larvae_month5, newdata = grid_vc_larvaepk2)
grid_vc_larvaepk2$se2 <- predict(pk_larvae_month5, newdata = grid_vc_larvaepk2, se = T)[[2]]
grid_vc_larvaepk2$pred2_up <- grid_vc_larvaepk2$pred2 + 1.96 * grid_vc_larvaepk2$se2
grid_vc_larvaepk2$pred2_lw <- grid_vc_larvaepk2$pred2 - 1.96 * grid_vc_larvaepk2$se2
```

Create two panel plot
Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 2), 
    mai = c(0.8, 0.9, 0.5, 0.5),
    family = "serif")
map_vc(pk_larvae, grid_vc_larvaepk, grid_vc_larvaepk2)
  legend(110,
         4,
         legend = c('Mean SST', '+ 1C'),
         col = c('#000000', '#E69F00'),
         lty = 1,
         bty = 'n',
         lwd = 2,
         cex = 0.6)
dev.copy(jpeg,
         here('results/pollock_hindcast',
         'pollock_larvae_vc.jpg'),
         height = 3.5,
         width = 8,
         res = 200,
         units = 'in')
dev.off()
```

### Tweedie
Base models
```{r}
# Tweedie
pk_egg_base_t <- gam(catch ~ factor(year) +
                       s(lon, lat) + 
                       s(doy),
                    data = pk_egg,
                    family = tw(link = 'log'),
                    method = 'REML')
summary(pk_egg_base_t)

pk_larvae_base_t <- gam(catch ~ factor(year) +
                       s(lon, lat) + 
                       s(doy),
                    data = pk_larvae,
                    family = tw(link = 'log'),
                    method = 'REML')
summary(pk_larvae_base_t)
```
## Eggs

Models with flexible phenology or geography 
Note variable coefficient by year for each
This takes forever to run, don't do unless necessary to recreate figure
```{r}
pk_egg_phenology_t <- gam(catch ~ factor(year) +
                            s(doy) +
                            s(lon, lat) +
                            s(doy, by = factor(year)),
                          data = pk_egg,
                          family = tw(link = 'log'),
                          method =  'REML')
summary(pk_egg_phenology_t)
AIC(pk_egg_phenology_t)
# AIC: 33996

pk_egg_geography_t <- gam(catch ~ factor(year) +
                            s(lon, lat) +
                            s(lon, lat, by = factor(year)) +
                            s(doy),
                          data = pk_egg,
                          family = tw(link = 'log'),
                          method = 'REML')
summary(pk_egg_geography_t)
AIC(pk_egg_geography_t)
# AIC: 33177
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_pkegg_phenology_t <- (summary(pk_egg_base_t)$scale - 
                      summary(pk_egg_phenology_t)$scale) / summary(pk_egg_base_t)$scale
ratio_pkegg_phenology_t #0.111

ratio_pkegg_geography_t <- (summary(pk_egg_base_t)$scale - 
                      summary(pk_egg_geography_t)$scale) / summary(pk_egg_base_t)$scale
ratio_pkegg_geography_t #0.310
```

Add ROMS mean temperatures Use a representative area with averaged values from Feb - April for each year This can then be matched to the egg/larvae data for use in variable coefficient by SST instead of year as above

Determine best GAM from 7 formulations (see function)
```{r}
pkegg_gams_t <- gam_selection(pk_egg, roms_temps)
AIC(pkegg_gams_t[[1]][[1]]) #33639
AIC(pkegg_gams_t[[1]][[2]]) #33634
AIC(pkegg_gams_t[[1]][[3]]) #33570
summary(pkegg_gams_t[[2]]) # best model is spatial VC, retain all smooth terms
AIC(pkegg_gams_t[[1]][[1]]) - AIC(pkegg_gams_t[[1]][[3]])
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(pkegg_gams_t[[1]][[2]], select = 1, main = "DOY")
plot(pkegg_gams_t[[1]][[2]], select = 3, main = "Temperature")
plot(pkegg_gams_t[[1]][[2]], select = 4, main = "Salinity")
plot(pkegg_gams_t[[1]][[2]], select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_egg_roms_monthvc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(pkegg_gams_t[[1]][[2]])
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_egg_roms_monthvc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(pkegg_gams_t[[1]][[3]], select = 1, main = "DOY")
plot(pkegg_gams_t[[1]][[3]], select = 3, main = "Temperature")
plot(pkegg_gams_t[[1]][[3]], select = 4, main = "Salinity")
plot(pkegg_gams_t[[1]][[3]], select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_egg_roms_spacevc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(pkegg_gams_t[[1]][[3]])
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_egg_roms_spacevc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_pkegg_month_t <- (summary(pk_egg_base_t)$scale - 
                         summary(pkegg_gams_t[[1]][[2]])$scale) / summary(pk_egg_base_t)$scale
var_ratio_pkegg_month_t #0.0418

var_ratio_pkegg_space_t <- (summary(pk_egg_base_t)$scale - 
                         summary(pkegg_gams_t[[1]][[3]])$scale) / summary(pk_egg_base_t)$scale
var_ratio_pkegg_space_t #0.0562
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
pkegg_grids_t <- threepanel_base_mse(pk_egg, pk_egg_base_t)
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(pk_egg, pkegg_grids_t, ratio_pkegg_geography_t,
                 var_ratio_pkegg_space_t,
                 ratio_pkegg_phenology_t,
                 var_ratio_pkegg_month_t)
dev.copy(jpeg,
         here('results/pollock_hindcast', 
              'pollock_egg_MSE_t.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
```{r}
pkegg_vcgrids_t <- vc_grids(pk_egg, pkegg_gams_t)
```

Create two panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc(pk_egg, pkegg_gams_t[[2]], pkegg_vcgrids_t, "Hindcast")
legend(50,
       6,
       legend = c('Mean SST', '+ 1C'),
       col = c('#000000', '#E69F00'),
       lty = 1,
       bty = 'n',
       lwd = 2,
       cex = 1.2)
dev.copy(jpeg,
         here('results/pollock_hindcast',
         'pollock_egg_vc_t_presentation.jpg'),
         height = 3.5,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

```{r}
tiff(here('results/pollock_hindcast/outputs',
         'pollock_egg_output.jpg'),
     units = "in",
     width = 40,
     height = 12,
     res = 200)

par(mfrow = c(1, 3),
    mar = c(11, 15, .5, 0.6) + 0.1,
    oma = c(3, 1, 1, 1),
    mgp = c(9, 4, 0))
plot_variable(pkegg_gams_t,
              covariate = 1,
              bounds = c(-5.5, 7),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(pkegg_gams_t,
              covariate = 3,
              bounds = c(-5.5, 7),
              "Temperature",
              " ",
              "n")
plot_variable(pkegg_gams_t,
              covariate = 4,
              bounds = c(-5.5, 7),
              "Salinity",
              " ",
              "n")

dev.off()
```


## Larvae

Models with flexible phenology or geography 
Note variable coefficient by year for each
```{r}
pk_larvae_phenology_t <- gam(catch ~ factor(year) +
                               s(doy) +
                               s(lon, lat) +
                               s(doy, by = factor(year)),
                             data = pk_larvae,
                             family = tw(link = 'log'),
                             method = 'REML')
summary(pk_larvae_phenology_t)
AIC(pk_larvae_phenology_t)
# AIC: 32891

pk_larvae_geography_t <- gam(catch ~ factor(year) +
                               s(lon, lat) +
                               s(lon, lat, by = factor(year)) +
                               s(doy),
                             data = pk_larvae,
                             family = tw(link = 'log'),
                             method = 'REML')
summary(pk_larvae_geography_t)
AIC(pk_larvae_geography_t)
# AIC: 31370
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_pklarvae_phenology_t <- (summary(pk_larvae_base_t)$scale - 
                      summary(pk_larvae_phenology_t)$scale) / summary(pk_larvae_base_t)$scale
ratio_pklarvae_phenology_t # 0.119

ratio_pklarvae_geography_t <- (summary(pk_larvae_base_t)$scale - 
                      summary(pk_larvae_geography_t)$scale) / summary(pk_larvae_base_t)$scale
ratio_pklarvae_geography_t # 0.426
```

Determine best GAM from 7 formulations (see function)
```{r}
pklarvae_gams_t <- gam_selection(pk_larvae, roms_temps)
AIC(pklarvae_gams_t[[1]][[1]]) #32523
AIC(pklarvae_gams_t[[1]][[2]]) #32497
AIC(pklarvae_gams_t[[1]][[3]]) #32421
summary(pklarvae_gams_t[[2]]) # best model is spatial, retain all smooth terms
AIC(pklarvae_gams_t[[1]][[1]]) - AIC(pklarvae_gams_t[[2]]) #102.0
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(pklarvae_gams_t[[1]][[2]], select = 1, main = "DOY")
plot(pklarvae_gams_t[[1]][[2]], select = 3, main = "Temperature")
plot(pklarvae_gams_t[[1]][[2]], select = 4, main = "Salinity")
plot(pklarvae_gams_t[[1]][[2]], select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_larvae_roms_monthvc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(pklarvae_gams_t[[1]][[2]])
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_larvae_roms_monthvc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(pklarvae_gams_t[[1]][[3]], select = 1, main = "DOY")
plot(pklarvae_gams_t[[1]][[3]], select = 3, main = "Temperature")
plot(pklarvae_gams_t[[1]][[3]], select = 4, main = "Salinity")
plot(pklarvae_gams_t[[1]][[3]], select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_larvae_roms_spacevc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(pklarvae_gams_t[[1]][[3]])
dev.copy(jpeg, here('results/pollock_hindcast', 'pollock_larvae_roms_spacevc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_pklarvae_month_t <- (summary(pk_larvae_base_t)$scale - 
                         summary(pklarvae_gams_t[[1]][[2]])$scale) / summary(pk_larvae_base_t)$scale
var_ratio_pklarvae_month_t #0.0737

var_ratio_pklarvae_space_t <- (summary(pk_larvae_base_t)$scale - 
                         summary(pklarvae_gams_t[[1]][[3]])$scale) / summary(pk_larvae_base_t)$scale
var_ratio_pklarvae_space_t #0.0900
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
pklarvae_grids_t <- threepanel_base_mse(pk_larvae, pk_larvae_base_t)
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(pk_larvae, pklarvae_grids_t, ratio_pklarvae_geography_t,
                 var_ratio_pklarvae_space_t,
                 ratio_pklarvae_phenology_t,
                 var_ratio_pklarvae_month_t)
dev.copy(jpeg,
         here('results/pollock_hindcast', 
              'pollock_larvae_MSE_t.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
```{r}
pklarvae_vcgrids_t <- vc_grids(pk_larvae, pklarvae_gams_t)
```

Create two panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc(pk_larvae, pklarvae_gams_t[[2]], pklarvae_vcgrids_t, "Spatial GAM hindcast")
legend(50,
       2,
       legend = c('Mean SST', '+ 1C'),
       col = c('#000000', '#E69F00'),
       lty = 1,
       bty = 'n',
       lwd = 2,
       cex = 1.2)
dev.copy(jpeg,
         here('results/pollock_hindcast',
         'pollock_larvae_vc_t.jpg'),
         height = 3.5,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

```{r}
tiff(here('results/pollock_hindcast/outputs',
         'pollock_larvae_output.jpg'),
     units = "in",
     width = 40,
     height = 12,
     res = 200)

par(mfrow = c(1, 3),
    mar = c(11, 15, .5, 0.6) + 0.1,
    oma = c(3, 1, 1, 1),
    mgp = c(9, 4, 0))
plot_variable(pklarvae_gams_t,
              covariate = 1,
              bounds = c(-5.5, 7),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(pklarvae_gams_t,
              covariate = 3,
              bounds = c(-5.5, 7),
              "Temperature",
              " ",
              "n")
plot_variable(pklarvae_gams_t,
              covariate = 4,
              bounds = c(-5.5, 7),
              "Salinity",
              " ",
              "n")

dev.off()
```

#### Flathead Sole
### Tweedie
```{r}
fhs_egg_base_t <- gam(catch ~ factor(year) +
                        s(lon, lat) +
                        s(doy),
                      data = fhs_egg,
                      family = tw(link = 'log'),
                      method = 'REML')
summary(fhs_egg_base_t)

fhs_larvae_base_t <- gam(catch ~ factor(year) +
                           s(lon, lat) +
                           s(doy),
                         data = fhs_larvae,
                         family = tw(link = 'log'),
                         method = 'REML')
summary(fhs_larvae_base_t)
```

## Eggs

Models with flexible phenology or geography 
Note variable coefficient by year for each
This takes forever to run, don't do unless necessary to recreate figure
```{r}
fhs_egg_phenology_t <- gam(catch ~ factor(year) +
                             s(doy) +
                             s(lon, lat) +
                             s(doy, by = factor(year)),
                           data = fhs_egg,
                           family = tw(link = 'log'),
                           method = 'REML')
summary(fhs_egg_phenology_t)
AIC(fhs_egg_phenology_t)
# AIC: 9907

fhs_egg_geography_t <- gam(catch ~ factor(year) +
                             s(lon, lat) +
                             s(lon, lat, by = factor(year)) +
                             s(doy),
                           data = fhs_egg,
                           family = tw(link = 'log'),
                           method =  'REML')
summary(fhs_egg_geography_t)
AIC(fhs_egg_geography_t)
# AIC: 8939
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_fhsegg_phenology_t <- (summary(fhs_egg_base_t)$scale - 
                      summary(fhs_egg_phenology_t)$scale) / summary(fhs_egg_base_t)$scale
ratio_fhsegg_phenology_t #0.171

ratio_fhsegg_geography_t <- (summary(fhs_egg_base_t)$scale - 
                      summary(fhs_egg_geography_t)$scale) / summary(fhs_egg_base_t)$scale
ratio_fhsegg_geography_t #0.496
```

Add ROMS mean temperatures Use a representative area with averaged values from Feb - April for each year This can then be matched to the egg/larvae data for use in variable coefficient by SST instead of year as above


Determine best GAM from 7 formulations (see function)
```{r}
fhsegg_gams_t <- gam_selection(fhs_egg, roms_temps)
AIC(fhsegg_gams_t[[1]][[1]]) # 13559
AIC(fhsegg_gams_t[[1]][[2]]) # 13343
AIC(fhsegg_gams_t[[1]][[3]]) # 13536
summary(fhsegg_gams_t[[2]]) # best model is doy VC, remove salinity
AIC(fhsegg_gams_t[[1]][[1]]) - AIC(fhsegg_gams_t[[2]]) #216.7
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(fhsegg_gams_t[[1]][[2]], select = 1, main = "DOY")
plot(fhsegg_gams_t[[1]][[2]], select = 3, main = "Temperature")
plot(fhsegg_gams_t[[1]][[2]], select = 4, main = "Salinity")
plot(fhsegg_gams_t[[1]][[2]], select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/flathead_hindcast', 'flathead_egg_roms_monthvc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(fhsegg_gams_t[[1]][[2]])
dev.copy(jpeg, here('results/flathead_hindcast', 'flathead_egg_roms_monthvc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(fhsegg_gams_t[[1]][[3]], select = 1, main = "DOY")
plot(fhsegg_gams_t[[1]][[3]], select = 3, main = "Temperature")
plot(fhsegg_gams_t[[1]][[3]], select = 4, main = "Salinity")
plot(fhsegg_gams_t[[1]][[3]], select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/flathead_hindcast', 'flathead_egg_roms_spacevc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(fhsegg_gams_t[[1]][[3]])
dev.copy(jpeg, here('results/flathead_hindcast', 'flathead_egg_roms_spacevc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_fhsegg_month_t <- (summary(fhs_egg_base_t)$scale - 
                         summary(fhsegg_gams_t[[1]][[2]])$scale) / summary(fhs_egg_base_t)$scale
var_ratio_fhsegg_month_t #0.230

var_ratio_fhsegg_space_t <- (summary(fhs_egg_base_t)$scale - 
                         summary(fhsegg_gams_t[[1]][[3]])$scale) / summary(fhs_egg_base_t)$scale
var_ratio_fhsegg_space_t #0.189
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
fhsegg_grids_t <- threepanel_base_mse(fhs_egg, fhs_egg_base_t)
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(fhs_egg, fhsegg_grids_t, ratio_fhsegg_geography_t,
                 var_ratio_fhsegg_space_t,
                 ratio_fhsegg_phenology_t,
                 var_ratio_fhsegg_month_t)
dev.copy(jpeg,
         here('results/flathead_hindcast', 
              'flathead_egg_MSE_t.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
```{r}
fhsegg_vcgrids_t <- vc_grids(fhs_egg, fhsegg_gams_t)
```

Create three panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc(fhs_egg, fhsegg_gams_t[[2]], fhsegg_vcgrids_t, "Spatial GAM hindcast")
legend(140,
       0,
       legend = c('Mean SST', '+ 1C'),
       col = c('#000000', '#E69F00'),
       lty = 1,
       bty = 'n',
       lwd = 2,
       cex = 1.2)
dev.copy(jpeg,
         here('results/flathead_hindcast',
         'flathead_egg_vc_t.jpg'),
         height = 3.5,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

```{r}
tiff(here('results/flathead_hindcast/outputs',
         'flathead_egg_output.jpg'),
     units = "in",
     width = 40,
     height = 12,
     res = 200)

par(mfrow = c(1, 3),
    mar = c(11, 15, .5, 0.6) + 0.1,
    oma = c(3, 1, 1, 1),
    mgp = c(9, 4, 0))
plot_variable(fhsegg_gams_t,
              covariate = 1,
              bounds = c(-5, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(fhsegg_gams_t,
              covariate = 3,
              bounds = c(-5, 3),
              "Temperature",
              " ",
              "n")
plot_variable(fhsegg_gams_t,
              covariate = 4,
              bounds = c(-5, 3),
              "Salinity",
              " ",
              "n")

dev.off()
```

## Larvae

Models with flexible phenology or geography 
Note variable coefficient by year for each
```{r}
fhs_larvae_phenology_t <- gam(catch ~ factor(year) +
                                s(doy) +
                                s(lon, lat) +
                                s(doy, by = factor(year)),
                              data = fhs_larvae,
                              family = tw(link = 'log'),
                              method = 'REML')
summary(fhs_larvae_phenology_t)
AIC(fhs_larvae_phenology_t)
# AIC: 22747

fhs_larvae_geography_t <- gam(catch ~ factor(year) +
                                s(lon, lat) +
                                s(lon, lat, by = factor(year)) +
                                s(doy),
                              data = fhs_larvae,
                              family = tw(link = 'log'),
                              method = 'REML')
summary(fhs_larvae_geography_t)
AIC(fhs_larvae_geography_t)
# AIC: 21956
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_fhslarvae_phenology_t <- (summary(fhs_larvae_base_t)$scale - 
                      summary(fhs_larvae_phenology_t)$scale) / summary(fhs_larvae_base_t)$scale
ratio_fhslarvae_phenology_t # 0.121

ratio_fhslarvae_geography_t <- (summary(fhs_larvae_base_t)$scale - 
                      summary(fhs_larvae_geography_t)$scale) / summary(fhs_larvae_base_t)$scale
ratio_fhslarvae_geography_t # 0.271
```

Add ROMS mean temperatures Use a representative area with averaged values from Feb - April for each year This can then be matched to the larvae/larvae data for use in variable coefficient by SST instead of year as above

Determine best GAM from 7 formulations (see function)
```{r}
fhslarvae_gams_t <- gam_selection(fhs_larvae, roms_temps)
AIC(fhslarvae_gams_t[[1]][[1]]) #14864
AIC(fhslarvae_gams_t[[1]][[2]]) #14847
AIC(fhslarvae_gams_t[[1]][[3]]) #14843
summary(fhslarvae_gams_t[[2]]) # best model is spatial VC, keep all smooth terms
AIC(fhslarvae_gams_t[[1]][[1]]) - AIC(fhslarvae_gams_t[[2]]) #20.2
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(fhslarvae_gams_t[[1]][[2]], select = 1, main = "DOY")
plot(fhslarvae_gams_t[[1]][[2]], select = 3, main = "Temperature")
plot(fhslarvae_gams_t[[1]][[2]], select = 4, main = "Salinity")
plot(fhslarvae_gams_t[[1]][[2]], select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/flathead_hindcast', 'flathead_larvae_roms_monthvc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(fhslarvae_gams_t[[1]][[2]])
dev.copy(jpeg, here('results/flathead_hindcast', 'flathead_larvae_roms_monthvc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(fhslarvae_gams_t[[1]][[3]], select = 1, main = "DOY")
plot(fhslarvae_gams_t[[1]][[3]], select = 3, main = "Temperature")
plot(fhslarvae_gams_t[[1]][[3]], select = 4, main = "Salinity")
plot(fhslarvae_gams_t[[1]][[3]], select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/flathead_hindcast', 'flathead_larvae_roms_spacevc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(fhslarvae_gams_t[[1]][[3]])
dev.copy(jpeg, here('results/flathead_hindcast', 'flathead_larvae_roms_spacevc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_fhslarvae_month_t <- (summary(fhs_larvae_base_t)$scale - 
                         summary(fhslarvae_gams_t[[1]][[2]])$scale) / summary(fhs_larvae_base_t)$scale
var_ratio_fhslarvae_month_t #0.0701

var_ratio_fhslarvae_space_t <- (summary(fhs_larvae_base_t)$scale - 
                         summary(fhslarvae_gams_t[[1]][[3]])$scale) / summary(fhs_larvae_base_t)$scale
var_ratio_fhslarvae_space_t #0.0725
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
fhslarvae_grids_t <- threepanel_base_mse(fhs_larvae, fhs_larvae_base_t)
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(fhs_larvae, fhslarvae_grids_t, ratio_fhslarvae_geography_t,
                 var_ratio_fhslarvae_space_t,
                 ratio_fhslarvae_phenology_t,
                 var_ratio_fhslarvae_month_t)
dev.copy(jpeg,
         here('results/flathead_hindcast', 
              'flathead_larvae_MSE_t.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
```{r}
fhslarvae_vcgrids_t <- vc_grids(fhs_larvae, fhslarvae_gams_t)
```

Create three panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc(fhs_larvae, fhslarvae_gams_t[[2]], fhslarvae_vcgrids_t, "Spatial GAM hindcast")
legend(40,
       0.5,
       legend = c('Mean SST', '+ 1C'),
       col = c('#000000', '#E69F00'),
       lty = 1,
       bty = 'n',
       lwd = 2,
       cex = 1.2)
dev.copy(jpeg,
         here('results/flathead_hindcast',
         'flathead_larvae_vc_t.jpg'),
         height = 3.5,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

```{r}
tiff(here('results/flathead_hindcast/outputs',
         'flathead_larvae_output.jpg'),
     units = "in",
     width = 40,
     height = 12,
     res = 200)

par(mfrow = c(1, 3),
    mar = c(11, 15, .5, 0.6) + 0.1,
    oma = c(3, 1, 1, 1),
    mgp = c(9, 4, 0))
plot_variable(fhslarvae_gams_t,
              covariate = 1,
              bounds = c(-5, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(fhslarvae_gams_t,
              covariate = 3,
              bounds = c(-5, 3),
              "Temperature",
              " ",
              "n")
plot_variable(fhslarvae_gams_t,
              covariate = 4,
              bounds = c(-5, 3),
              "Salinity",
              " ",
              "n")
dev.off()
```

#### Yellowfin Sole
### Tweedie
```{r}
yfs_egg_base_t <- gam(catch ~ factor(year) +
                        s(lon, lat) +
                        s(doy),
                      data = yfs_egg,
                      family = tw(link = 'log'),
                      method = 'REML')
summary(yfs_egg_base_t)

yfs_larvae_base_t <- gam(catch ~ factor(year) +
                           s(lon, lat) +
                           s(doy),
                         data = yfs_larvae,
                         family = tw(link = 'log'),
                         method = 'REML')
summary(yfs_larvae_base_t)
```

## Eggs

Models with flexible phenology or geography 
Note variable coefficient by year for each
This takes forever to run, don't do unless necessary to recreate figure
```{r}
yfs_egg_phenology_t <- gam(catch ~ factor(year) +
                             s(doy) +
                             s(lon, lat) +
                             s(doy, by = factor(year)),
                           data = yfs_egg,
                           family = tw(link = 'log'),
                           method = 'REML')
summary(yfs_egg_phenology_t)
AIC(yfs_egg_phenology_t)
# AIC: 1586

yfs_egg_geography_t <- gam(catch ~ factor(year) +
                             s(lon, lat) +
                             s(lon, lat, by = factor(year)) +
                             s(doy),
                           data = yfs_egg,
                           family = tw(link = 'log'),
                           method =  'REML')
summary(yfs_egg_geography_t)
AIC(yfs_egg_geography_t)
# AIC: 997
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_yfsegg_phenology_t <- (summary(yfs_egg_base_t)$scale - 
                      summary(yfs_egg_phenology_t)$scale) / summary(yfs_egg_base_t)$scale
ratio_yfsegg_phenology_t # 0.250

ratio_yfsegg_geography_t <- (summary(yfs_egg_base_t)$scale - 
                      summary(yfs_egg_geography_t)$scale) / summary(yfs_egg_base_t)$scale
ratio_yfsegg_geography_t #0.620
```

Add ROMS mean temperatures Use a representative area with averaged values from Feb - April for each year This can then be matched to the egg/larvae data for use in variable coefficient by SST instead of year as above


Determine best GAM from 7 formulations (see function)
```{r}
yfsegg_gams_t <- gam_selection(yfs_egg, roms_temps)
AIC(yfsegg_gams_t[[1]][[1]]) # 1830
AIC(yfsegg_gams_t[[1]][[2]]) # 1774
AIC(yfsegg_gams_t[[1]][[3]]) # 1774
summary(yfsegg_gams_t[[2]]) # best model is base model, use VC spatial, no environmental smooth terms significant
AIC(yfsegg_gams_t[[1]][[2]]) - AIC(yfsegg_gams_t[[2]]) #0
```

Plot best model for phenology
```{r}
# Best model by AIC is model 1
par(mfrow = c(2, 2))
plot(yfsegg_gams_t[[1]][[2]], select = 1, main = "DOY")
plot(yfsegg_gams_t[[1]][[2]], select = 3, main = "Temperature")
plot(yfsegg_gams_t[[1]][[2]], select = 4, main = "Salinity")
plot(yfsegg_gams_t[[1]][[2]], select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/yellowfin_hindcast', 'yellowfin_egg_roms_monthvc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(yfsegg_gams_t[[1]][[2]])
dev.copy(jpeg, here('results/yellowfin_hindcast', 'yellowfin_egg_roms_monthvc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(yfsegg_gams_t[[1]][[3]], select = 1, main = "DOY")
plot(yfsegg_gams_t[[1]][[3]], select = 3, main = "Temperature")
plot(yfsegg_gams_t[[1]][[3]], select = 4, main = "Salinity")
plot(yfsegg_gams_t[[1]][[3]], select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/yellowfin_hindcast', 'yellowfin_egg_roms_spacevc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(yfsegg_gams_t[[1]][[3]])
dev.copy(jpeg, here('results/yellowfin_hindcast', 'yellowfin_egg_roms_spacevc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_yfsegg_month_t <- (summary(yfs_egg_base_t)$scale - 
                         summary(yfsegg_gams_t[[1]][[2]])$scale) / summary(yfs_egg_base_t)$scale
var_ratio_yfsegg_month_t #-0.0322

var_ratio_yfsegg_space_t <- (summary(yfs_egg_base_t)$scale - 
                         summary(yfsegg_gams_t[[1]][[3]])$scale) / summary(yfs_egg_base_t)$scale
var_ratio_yfsegg_space_t #-0.0282
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
yfsegg_grids_t <- threepanel_base_mse(yfs_egg, yfs_egg_base_t)
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(yfs_egg, yfsegg_grids_t, ratio_yfsegg_geography_t,
                 var_ratio_yfsegg_space_t,
                 ratio_yfsegg_phenology_t,
                 var_ratio_yfsegg_month_t)
dev.copy(jpeg,
         here('results/yellowfin_hindcast', 
              'yellowfin_egg_MSE_t.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
```{r}
yfsegg_vcgrids_t <- vc_grids(yfs_egg, yfsegg_gams_t)
```

Create three panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc(yfs_egg, yfsegg_gams_t[[1]][[2]], yfsegg_vcgrids_t, "Spatial GAM hindcast")
legend(40,
       2,
       legend = c('Mean SST', '+ 1C'),
       col = c('#000000', '#E69F00'),
       lty = 1,
       bty = 'n',
       lwd = 2,
       cex = 1.2)
dev.copy(jpeg,
         here('results/yellowfin_hindcast',
         'yellowfin_egg_vc_t.jpg'),
         height = 3.5,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

```{r}
tiff(here('results/yellowfin_hindcast/outputs',
         'yellowfin_egg_output.jpg'),
     units = "in",
     width = 40,
     height = 12,
     res = 200)

par(mfrow = c(1, 3),
    mar = c(11, 15, .5, 0.6) + 0.1,
    oma = c(3, 1, 1, 1),
    mgp = c(9, 4, 0))
plot_variable(yfsegg_gams_t,
              covariate = 1,
              bounds = c(-20, 10),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(yfsegg_gams_t,
              covariate = 3,
              bounds = c(-20, 10),
              "Temperature",
              " ",
              "n")
plot_variable(yfsegg_gams_t,
              covariate = 4,
              bounds = c(-20, 10),
              "Salinity",
              " ",
              "n")

dev.off()
```

## Larvae

Models with flexible phenology or geography 
Note variable coefficient by year for each
```{r}
yfs_larvae_phenology_t <- gam(catch ~ factor(year) +
                                s(doy) +
                                s(lon, lat) +
                                s(doy, by = factor(year)),
                              data = yfs_larvae,
                              family = tw(link = 'log'),
                              method =  'REML')
summary(yfs_larvae_phenology_t)
AIC(yfs_larvae_phenology_t)
# AIC: 8269

yfs_larvae_geography_t <- gam(catch ~ factor(year) +
                                s(lon, lat) +
                                s(lon, lat, by = factor(year)) +
                                s(doy),
                              data = yfs_larvae,
                              family = tw(link = 'log'),
                              method = 'REML')
summary(yfs_larvae_geography_t)
AIC(yfs_larvae_geography_t)
# AIC: 7970
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_yfslarvae_phenology_t <- (summary(yfs_larvae_base_t)$scale - 
                      summary(yfs_larvae_phenology_t)$scale) / summary(yfs_larvae_base_t)$scale
ratio_yfslarvae_phenology_t # 0.166

ratio_yfslarvae_geography_t <- (summary(yfs_larvae_base_t)$scale - 
                      summary(yfs_larvae_geography_t)$scale) / summary(yfs_larvae_base_t)$scale
ratio_yfslarvae_geography_t # 0.371
```

Add ROMS mean temperatures Use a representative area with averaged values from Feb - April for each year This can then be matched to the larvae/larvae data for use in variable coefficient by SST instead of year as above

Determine best GAM from 7 formulations (see function)
```{r}
yfslarvae_gams_t <- gam_selection(yfs_larvae, roms_temps)
AIC(yfslarvae_gams_t[[1]][[1]]) # 7932
AIC(yfslarvae_gams_t[[1]][[2]]) # 7891
AIC(yfslarvae_gams_t[[1]][[3]]) # 7852
summary(yfslarvae_gams_t[[2]]) # best is spatial VC, keep all smooth terms
AIC(yfslarvae_gams_t[[1]][[1]]) - AIC(yfslarvae_gams_t[[2]]) # 80.3
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(yfslarvae_gams_t[[1]][[2]], select = 1, main = "DOY")
plot(yfslarvae_gams_t[[1]][[2]], select = 3, main = "Temperature")
plot(yfslarvae_gams_t[[1]][[2]], select = 4, main = "Salinity")
plot(yfslarvae_gams_t[[1]][[2]], select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/yellowfin_hindcast', 'yellowfin_larvae_roms_monthvc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(yfslarvae_gams_t[[1]][[2]])
dev.copy(jpeg, here('results/yellowfin_hindcast', 'yellowfin_larvae_roms_monthvc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(yfslarvae_gams_t[[1]][[3]], select = 1, main = "DOY")
plot(yfslarvae_gams_t[[1]][[3]], select = 3, main = "Temperature")
plot(yfslarvae_gams_t[[1]][[3]], select = 4, main = "Salinity")
plot(yfslarvae_gams_t[[1]][[3]], select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/yellowfin_hindcast', 'yellowfin_larvae_roms_spacevc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(yfslarvae_gams_t[[1]][[3]])
dev.copy(jpeg, here('results/yellowfin_hindcast', 'yellowfin_larvae_roms_spacevc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_yfslarvae_month_t <- (summary(yfs_larvae_base_t)$scale - 
                         summary(yfslarvae_gams_t[[1]][[2]])$scale) / summary(yfs_larvae_base_t)$scale
var_ratio_yfslarvae_month_t #0.078

var_ratio_yfslarvae_space_t <- (summary(yfs_larvae_base_t)$scale - 
                         summary(yfslarvae_gams_t[[1]][[3]])$scale) / summary(yfs_larvae_base_t)$scale
var_ratio_yfslarvae_space_t #0.956
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
yfslarvae_grids_t <- threepanel_base_mse(yfs_larvae, yfs_larvae_base_t)
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(yfs_larvae, yfslarvae_grids_t, ratio_yfslarvae_geography_t,
                 var_ratio_yfslarvae_space_t,
                 ratio_yfslarvae_phenology_t,
                 var_ratio_yfslarvae_month_t)
dev.copy(jpeg,
         here('results/yellowfin_hindcast', 
              'yellowfin_larvae_MSE_t.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
```{r}
yfslarvae_vcgrids_t <- vc_grids(yfs_larvae, yfslarvae_gams_t)
```

Create three panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc(yfs_larvae, yfslarvae_gams_t[[2]], yfslarvae_vcgrids_t, "Spatial GAM hindcast")
legend(140,
       21,
       legend = c('Mean SST', '+ 1C'),
       col = c('#000000', '#E69F00'),
       lty = 1,
       bty = 'n',
       lwd = 2,
       cex = 1.2)
dev.copy(jpeg,
         here('results/yellowfin_hindcast',
         'yellowfin_larvae_vc_t.jpg'),
         height = 3.5,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

```{r}
tiff(here('results/yellowfin_hindcast/outputs',
         'yellowfin_larvae_output.jpg'),
     units = "in",
     width = 40,
     height = 12,
     res = 200)

par(mfrow = c(1, 3),
    mar = c(11, 15, .5, 0.6) + 0.1,
    oma = c(3, 1, 1, 1),
    mgp = c(9, 4, 0))
plot_variable(yfslarvae_gams_t,
              covariate = 1,
              bounds = c(-20, 10),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(yfslarvae_gams_t,
              covariate = 3,
              bounds = c(-20, 10),
              "Temperature",
              " ",
              "n")
plot_variable(yfslarvae_gams_t,
              covariate = 4,
              bounds = c(-20, 10),
              "Salinity",
              " ",
              "n")

dev.off()
```

#### Alaska Plaice
### Tweedie
```{r}
akp_egg_base_t <- gam(catch ~ factor(year) +
                        s(lon, lat) +
                        s(doy),
                      data = akp_egg,
                      family = tw(link = 'log'),
                      method =  'REML')
summary(akp_egg_base_t)

akp_larvae_base_t <- gam(catch ~ factor(year) +
                           s(lon, lat) +
                           s(doy),
                         data = akp_larvae,
                         family = tw(link = 'log'),
                         method = 'REML')
summary(akp_larvae_base_t)
```

## Eggs

Models with flexible phenology or geography 
Note variable coefficient by year for each
This takes forever to run, don't do unless necessary to recreate figure
```{r}
akp_egg_phenology_t <- gam(catch ~ factor(year) +
                             s(doy) +
                             s(lon, lat) +
                             s(doy, by = factor(year)),
                           data = akp_egg,
                           family = tw(link = 'log'),
                           method =  'REML')
summary(akp_egg_phenology_t)
AIC(akp_egg_phenology_t)
# AIC: 13892

akp_egg_geography_t <- gam(catch ~ factor(year) +
                             s(lon, lat) +
                             s(lon, lat, by = factor(year)) +
                             s(doy),
                           data = akp_egg,
                           family = tw(link = 'log'),
                           method =  'REML')
summary(akp_egg_geography_t)
AIC(akp_egg_geography_t)
# AIC: 13005
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_akpegg_phenology_t <- (summary(akp_egg_base_t)$scale - 
                      summary(akp_egg_phenology_t)$scale) / summary(akp_egg_base_t)$scale
ratio_akpegg_phenology_t # 0.195

ratio_akpegg_geography_t <- (summary(akp_egg_base_t)$scale - 
                      summary(akp_egg_geography_t)$scale) / summary(akp_egg_base_t)$scale
ratio_akpegg_geography_t # 0.443
```

Add ROMS mean temperatures Use a representative area with averaged values from Feb - April for each year This can then be matched to the egg/larvae data for use in variable coefficient by SST instead of year as above


Determine best GAM from 7 formulations (see function)
```{r}
akpegg_gams_t <- gam_selection(akp_egg, roms_temps)
AIC(akpegg_gams_t[[1]][[1]]) # 14191
AIC(akpegg_gams_t[[1]][[2]]) # 14140
AIC(akpegg_gams_t[[1]][[3]]) # 14143
summary(akpegg_gams_t[[2]]) # best model is temporal VC, keep all terms
AIC(akpegg_gams_t[[1]][[1]]) - AIC(akpegg_gams_t[[2]]) #50.4
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(akpegg_gams_t[[1]][[2]], select = 1, main = "DOY")
plot(akpegg_gams_t[[1]][[2]], select = 3, main = "Temperature")
plot(akpegg_gams_t[[1]][[2]], select = 4, main = "Salinity")
plot(akpegg_gams_t[[1]][[2]], select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/plaice_hindcast', 'plaice_egg_roms_monthvc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(akpegg_gams_t[[1]][[2]])
dev.copy(jpeg, here('results/plaice_hindcast', 'plaice_egg_roms_monthvc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(akpegg_gams_t[[1]][[3]], select = 1, main = "DOY")
plot(akpegg_gams_t[[1]][[3]], select = 3, main = "Temperature")
plot(akpegg_gams_t[[1]][[3]], select = 4, main = "Salinity")
plot(akpegg_gams_t[[1]][[3]], select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/plaice_hindcast', 'plaice_egg_roms_spacevc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(akpegg_gams_t[[1]][[3]])
dev.copy(jpeg, here('results/plaice_hindcast', 'plaice_egg_roms_spacevc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_akpegg_month_t <- (summary(akp_egg_base_t)$scale - 
                         summary(akpegg_gams_t[[1]][[2]])$scale) / summary(akp_egg_base_t)$scale
var_ratio_akpegg_month_t # 0.122

var_ratio_akpegg_space_t <- (summary(akp_egg_base_t)$scale - 
                         summary(akpegg_gams_t[[1]][[3]])$scale) / summary(akp_egg_base_t)$scale
var_ratio_akpegg_space_t # 0.111
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
akpegg_grids_t <- threepanel_base_mse(akp_egg, akp_egg_base_t)
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(akp_egg, akpegg_grids_t, ratio_akpegg_geography_t,
                 var_ratio_akpegg_space_t,
                 ratio_akpegg_phenology_t,
                 var_ratio_akpegg_month_t)
dev.copy(jpeg,
         here('results/plaice_hindcast', 
              'plaice_egg_MSE_t.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
```{r}
akpegg_vcgrids_t <- vc_grids(akp_egg, akpegg_gams_t)
```

Create three panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc(akp_egg, akpegg_gams_t[[2]], akpegg_vcgrids_t, "Spatial GAM hindcast")
legend(120,
       5,
       legend = c('Mean SST', '+ 1C'),
       col = c('#000000', '#E69F00'),
       lty = 1,
       bty = 'n',
       lwd = 2,
       cex = 1.2)
dev.copy(jpeg,
         here('results/plaice_hindcast',
         'plaice_egg_vc_t.jpg'),
         height = 3.5,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

```{r}
tiff(here('results/plaice_hindcast/outputs',
         'plaice_egg_output.jpg'),
     units = "in",
     width = 52,
     height = 12,
     res = 200)

par(mfrow = c(1, 4),
    mar = c(11, 15, .5, 0.6) + 0.1,
    oma = c(3, 1, 1, 1),
    mgp = c(9, 4, 0))
plot_variable(akpegg_gams_t,
              covariate = 1,
              bounds = c(-7, 8),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(akpegg_gams_t,
              covariate = 3,
              bounds = c(-7, 8),
              "Temperature",
              " ",
              "n")
plot_variable(akpegg_gams_t,
              covariate = 4,
              bounds = c(-7, 8),
              "Salinity",
              " ",
              "n")
plot_variable(akpegg_gams_t,
              covariate = 5,
              bounds = c(-7, 8),
              "Day of Year",
              " ",
              "n")
dev.off()
```

## Larvae

Models with flexible phenology or geography 
Note variable coefficient by year for each
```{r}
akp_larvae_phenology_t <- gam(catch ~ factor(year) +
                                s(doy) +
                                s(lon, lat) +
                                s(doy, by = factor(year)),
                              data = akp_larvae,
                              family = tw(link = 'log'),
                              method = 'REML')
summary(akp_larvae_phenology_t)
AIC(akp_larvae_phenology_t)
# AIC: 6035

akp_larvae_geography_t <- gam(catch ~ factor(year) +
                                s(lon, lat) +
                                s(lon, lat, by = factor(year)) +
                                s(doy),
                              data = akp_larvae,
                              family = tw(link = 'log'),
                              method = 'REML')
summary(akp_larvae_geography_t)
AIC(akp_larvae_geography_t)
# AIC: 5071
```

Ratio for comparison between base and flexible models 
Use to look at mean square error, base should have higher error
```{r}
ratio_akplarvae_phenology_t <- (summary(akp_larvae_base_t)$scale - 
                      summary(akp_larvae_phenology_t)$scale) / summary(akp_larvae_base_t)$scale
ratio_akplarvae_phenology_t # 0.132

ratio_akplarvae_geography_t <- (summary(akp_larvae_base_t)$scale - 
                      summary(akp_larvae_geography_t)$scale) / summary(akp_larvae_base_t)$scale
ratio_akplarvae_geography_t # 0.554
```

Add ROMS mean temperatures Use a representative area with averaged values from Feb - April for each year This can then be matched to the larvae/larvae data for use in variable coefficient by SST instead of year as above

Determine best GAM from 7 formulations (see function)
```{r}
akplarvae_gams_t <- gam_selection(akp_larvae, roms_temps)
AIC(akplarvae_gams_t[[1]][[1]]) # 5768
AIC(akplarvae_gams_t[[1]][[2]]) # 5654
AIC(akplarvae_gams_t[[1]][[3]]) # 5660
summary(akplarvae_gams_t[[2]]) # best is phenology VC
AIC(akplarvae_gams_t[[1]][[1]]) - AIC(akplarvae_gams_t[[2]]) #114.1
```

Plot best model for phenology
```{r}
# Best model by AIC is model 2
par(mfrow = c(2, 2))
plot(akplarvae_gams_t[[1]][[2]], select = 1, main = "DOY")
plot(akplarvae_gams_t[[1]][[2]], select = 3, main = "Temperature")
plot(akplarvae_gams_t[[1]][[2]], select = 4, main = "Salinity")
plot(akplarvae_gams_t[[1]][[2]], select = 5, main = "DOY by Mean ROMS SST")
dev.copy(jpeg, here('results/plaice_hindcast', 'plaice_larvae_roms_monthvc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(akplarvae_gams_t[[1]][[2]])
dev.copy(jpeg, here('results/plaice_hindcast', 'plaice_larvae_roms_monthvc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

Plot best model for geography
```{r}
par(mfrow = c(2, 2))
plot(akplarvae_gams_t[[1]][[3]], select = 1, main = "DOY")
plot(akplarvae_gams_t[[1]][[3]], select = 3, main = "Temperature")
plot(akplarvae_gams_t[[1]][[3]], select = 4, main = "Salinity")
plot(akplarvae_gams_t[[1]][[3]], select = 5, main = "Location by Mean ROMS SST")
dev.copy(jpeg, here('results/plaice_hindcast', 'plaice_larvae_roms_spacevc_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()

par(mfrow = c(2, 2))
gam.check(akplarvae_gams_t[[1]][[3]])
dev.copy(jpeg, here('results/plaice_hindcast', 'plaice_larvae_roms_spacevc_resid_t.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

VC GAM ratios
```{r}
# Use the original presence model from before
var_ratio_akplarvae_month_t <- (summary(akp_larvae_base_t)$scale - 
                         summary(akplarvae_gams_t[[1]][[2]])$scale) / summary(akp_larvae_base_t)$scale
var_ratio_akplarvae_month_t # 0.169

var_ratio_akplarvae_space_t <- (summary(akp_larvae_base_t)$scale - 
                         summary(akplarvae_gams_t[[1]][[3]])$scale) / summary(akp_larvae_base_t)$scale
var_ratio_akplarvae_space_t # 0.166
```

Plot the results of the GAMs that vary by year This generates a three panel figure with a map, phenology curve, and MSE barplot Grids:
```{r}
akplarvae_grids_t <- threepanel_base_mse(akp_larvae, akp_larvae_base_t)
```

Plots:
```{r}
# Map & curve
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_phenology(akp_larvae, akplarvae_grids_t, ratio_akplarvae_geography_t,
                 var_ratio_akplarvae_space_t,
                 ratio_akplarvae_phenology_t,
                 var_ratio_akplarvae_month_t)
dev.copy(jpeg,
         here('results/plaice_hindcast', 
              'plaice_larvae_MSE_t.jpg'),
         height = 3.5,
         width = 12,
         res = 200,
         units = 'in')
```

Fill grids with output from SST VC GAMs
```{r}
akplarvae_vcgrids_t <- vc_grids(akp_larvae, akplarvae_gams_t)
```

Create three panel plot Map of distribution & phenology curves with +1 temperature change
```{r}
par(mfrow = c(1, 3), 
    mai = c(0.7, 0.6, 0.4, 0.4),
    family = "serif")
map_vc(akp_larvae, akplarvae_gams_t[[2]], akplarvae_vcgrids_t, "Spatial GAM hindcast")
legend(140,
       8,
       legend = c('Mean SST', '+ 1C'),
       col = c('#000000', '#E69F00'),
       lty = 1,
       bty = 'n',
       lwd = 2,
       cex = 1.2)
dev.copy(jpeg,
         here('results/plaice_hindcast',
         'plaice_larvae_vc_t.jpg'),
         height = 3.5,
         width = 10,
         res = 200,
         units = 'in')
dev.off()
```

```{r}
tiff(here('results/plaice_hindcast/outputs',
         'plaice_larvae_output.jpg'),
     units = "in",
     width = 52,
     height = 12,
     res = 200)

par(mfrow = c(1, 4),
    mar = c(11, 15, .5, 0.6) + 0.1,
    oma = c(3, 1, 1, 1),
    mgp = c(9, 4, 0))
plot_variable(akplarvae_gams_t,
              covariate = 1,
              bounds = c(-7, 8),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(akplarvae_gams_t,
              covariate = 3,
              bounds = c(-7, 8),
              "Temperature",
              " ",
              "n")
plot_variable(akplarvae_gams_t,
              covariate = 4,
              bounds = c(-7, 8),
              "Salinity",
              " ",
              "n")
plot_variable(akplarvae_gams_t,
              covariate = 5,
              bounds = c(-7, 8),
              "Day of Year",
              " ",
              "n")
dev.off()
```

#### GAM multipanel Figure
```{r}
#### GAM multipanel Figure
tiff(here('results',
          'hindcast_output.jpg'),
     units = "in",
     width = 65,
     height = 90,
     res = 300)
par(mfrow = c(7, 5),
    mar = c(11, 15, 5, 1.5) + 0.1,
    oma = c(3, 35, 15, 1),
    mgp = c(10, 4, 0),
    family = "serif")
plot_variable(pkegg_gams_t,
              covariate = 1,
              bounds = c(-5.5, 7),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(pkegg_gams_t,
              covariate = 3,
              bounds = c(-5.5, 7),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(pkegg_gams_t,
              covariate = 4,
              bounds = c(-5.5, 7),
              "Salinity",
              " ",
              "n")
plot(0, xaxt = "n", yaxt = "n", bty = "n", pch = "", ylab = "", xlab = "")
location_plot(pkegg_gams_t,
              pk_egg)
plot_variable(pklarvae_gams_t,
              covariate = 1,
              bounds = c(-5.5, 7),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(pklarvae_gams_t,
              covariate = 3,
              bounds = c(-5.5, 7),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(pklarvae_gams_t,
              covariate = 4,
              bounds = c(-5.5, 7),
              "Salinity",
              " ",
              "n")
plot(0, xaxt = "n", yaxt = "n", bty = "n", pch = "", ylab = "", xlab = "")
location_plot(pklarvae_gams_t,
              pk_larvae)
plot_variable(fhsegg_gams_t,
              covariate = 1,
              bounds = c(-6, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(fhsegg_gams_t,
              covariate = 3,
              bounds = c(-6, 3),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(fhsegg_gams_t,
              covariate = 4,
              bounds = c(-6, 3),
              "Salinity",
              " ",
              "n")
plot_variable(fhsegg_gams_t,
              covariate = 5,
              bounds = c(-6, 3),
              "Day of Year",
              " ",
              "n")
plot(0, xaxt = "n", yaxt = "n", bty = "n", pch = "", ylab = "", xlab = "")
plot_variable(fhslarvae_gams_t,
              covariate = 1,
              bounds = c(-6, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(fhslarvae_gams_t,
              covariate = 3,
              bounds = c(-6, 3),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(fhslarvae_gams_t,
              covariate = 4,
              bounds = c(-6, 3),
              "Salinity",
              " ",
              "n")
plot(0, xaxt = "n", yaxt = "n", bty = "n", pch = "", ylab = "", xlab = "")
location_plot(fhslarvae_gams_t,
              fhs_larvae)
plot_variable(akpegg_gams_t,
              covariate = 1,
              bounds = c(-7, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(akpegg_gams_t,
              covariate = 3,
              bounds = c(-7, 3),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(akpegg_gams_t,
              covariate = 4,
              bounds = c(-7, 3),
              "Salinity",
              " ",
              "n")
plot_variable(akpegg_gams_t,
              covariate = 5,
              bounds = c(-7, 3),
              "Day of Year",
              " ",
              "n")
plot(0, xaxt = "n", yaxt = "n", bty = "n", pch = "", ylab = "", xlab = "")
plot_variable(akplarvae_gams_t,
              covariate = 1,
              bounds = c(-7, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(akplarvae_gams_t,
              covariate = 3,
              bounds = c(-7, 3),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(akplarvae_gams_t,
              covariate = 4,
              bounds = c(-7, 3),
              "Salinity",
              " ",
              "n")
plot_variable(akplarvae_gams_t,
              covariate = 5,
              bounds = c(-7, 3),
              "Day of Year",
              " ",
              "n")
plot(0, xaxt = "n", yaxt = "n", bty = "n", pch = "", ylab = "", xlab = "")
plot_variable(yfslarvae_gams_t,
              covariate = 1,
              bounds = c(-4, 3),
              "Day of Year",
              "Species Abundance Anomalies",
              "s")
plot_variable(yfslarvae_gams_t,
              covariate = 3,
              bounds = c(-4, 3),
              expression(paste("Temperature ("^0, 'C)')),
              " ",
              "n")
plot_variable(yfslarvae_gams_t,
              covariate = 4,
              bounds = c(-4, 3),
              "Salinity",
              " ",
              "n")
plot(0, xaxt = "n", yaxt = "n", bty = "n", pch = "", ylab = "", xlab = "")
location_plot(yfslarvae_gams_t,
              yfs_larvae)
mtext("Walleye Pollock", 
      side = 2, 
      line = 25, 
      outer = TRUE, 
      cex = 7,
      at = 0.86)
mtext("Flathead Sole", 
      side = 2, 
      line = 25, 
      outer = TRUE, 
      cex = 7,
      at = 0.58)
mtext("Alaska Plaice", 
      side = 2, 
      line = 25, 
      outer = TRUE, 
      cex = 7,
      at = 0.29)
mtext("Yellowfin Sole", 
      side = 2, 
      line = 25, 
      outer = TRUE, 
      cex = 7,
      at = 0.08)
mtext("Day of Year", 
      side = 3, 
      line = 2, 
      outer = TRUE, 
      cex = 7,
      at = 0.11)
mtext("Temperature", 
      side = 3, 
      line = 2, 
      outer = TRUE, 
      cex = 7,
      at = 0.31)
mtext("Salinity", 
      side = 3, 
      line = 2, 
      outer = TRUE, 
      cex = 7,
      at = 0.51)
mtext("Day of Year \n by Mean SST",
      side = 3,
      line = -1.8,
      outer = TRUE,
      cex = 7,
      at = 0.71)
mtext("Location",
      side = 3,
      line = 2,
      outer = TRUE,
      cex = 7,
      at = 0.91)
mtext("Eggs", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.93)
mtext("Larvae", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.79)
mtext("Eggs", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.65)
mtext("Larvae", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.51)
mtext("Eggs", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.36)
mtext("Larvae", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.22)
mtext("Larvae", 
      side = 2, 
      line = 10, 
      outer = TRUE, 
      cex = 6,
      at = 0.08)
dev.off()
```